---
title: "Extracting Grid PM~2.5~ Values for Colorado"
author: "Ryan Gan"
date: "2018-03-23"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Introduction

I have a 3x3 km resolution grid I think for the entire state of Colorado, where I also have an ID that can joined to health outcomes data. I have exposure estimates at a larger grid size of 15x15 km. I need to regrid the NARR temperature estimates to match the Colorado grids.

I will need tidyverse, ncdf4, and raster.
```{r libraries}
library(tidyverse)
library(ncdf4)
library(raster)
```

## PM~2.5~ Grid Coordinates

Opening connection to 2015 PM~2.5~ nc file.

```{r pm_nc_open}
# open connection pm grid
pm_nc_path <- paste0("../../../meta_wildfire/data/smoke/",
  "krigedPM25_06-15/krigedPM25_2015.nc")
pm_nc <- nc_open(pm_nc_path)
```

Reading in Colorado grid sf file to get out sorted GRID_ID values to join with PM~2.5~ grid coordinates.

```{r colorado_grid_id}
# grid id from sf file
# define grid path
grid_path <- paste0("../../data/shapefiles/co_krig_grid/")
# read colorado grid
co_grid_sf <- sf::st_read(dsn = grid_path, layer = "co_krig_grid") %>% 
  arrange(ROWY, COLX)
# output grid id
GRID_ID <- co_grid_sf$GRID_ID
```

Extracting the Colorado grid coordinates to regrid temperature to and combine with GRID ID vector in a dataframe.

```{r colorado_grid_loc}
# extract latitude and longitude values
lat <- as.vector(ncvar_get(nc = pm_nc, varid = "lat")[104:147, 81:112])
lon <- as.vector(ncvar_get(nc = pm_nc, varid = "lon")[104:147, 81:112])
# bind with grid id and create dataframe
grid_coords <- data.frame(GRID_ID, lon, lat)
```

## Temperature Regrid

Opening up connection to the 2015 air temp at 2 meters file to get the coordinates.

```{r open_temp_nc}
# open temp nc
temp_nc <- nc_open(paste0("../../../meta_wildfire/data/smoke/",
  "air_temp_2m/air.2m.2015.nc"))
temp_nc
```

Extracting latitude and longitude coordinates for temperature grid.

```{r temp_coords}
# extract temp lat lon
t_lat <- as.vector(ncvar_get(nc = temp_nc, var = "lat"))
t_lon <- as.vector(ncvar_get(nc = temp_nc, var = "lon"))
# temp_coordinates as matrix
temp_coords <- as.matrix(cbind(t_lon, t_lat), ncol =2)
# set names to lat and lon
colnames(temp_coords) <- c("lon", "lat")
# view first obs of temp coordinates
head(temp_coords)
```

Now that I have the temperature coordinates, I'm going to extract the spatial extent of the coordinates and print the extent object boundaries.

```{r temp_extent}
# create extent object
e <- extent(temp_coords[,1:2])
# print extent object
e
```

I'm going to create an empty raster that matches the number of columns and rows in the temperature grid (349 columns and 277 rows).

```{r new_temp_raster}
# create new raster
r <- raster(e, nrow=277, ncol=349)
```

### Checking January 1st 2015 Temperature for Colorado

I'm going to extract the temperature values on January 1st 2015 to the Colorado grids to make sure it looks right.

```{r temp_vector}
# extract vector of temp
t_temp <- as.vector(ncvar_get(temp_nc, varid = "air")[,,1]) 
```

Rasterizing vector of January 1st temp to the new raster. Using mean for multiple values that fall in to the same grid cells for the new grid.

```{r rasterize}
# rasterize temperature values to new grid
r_temp <- rasterize(temp_coords[,1:2], r, t_temp, fun=mean)
# extract values to pm points
temp_k_regrid <- extract(r_temp, grid_coords[,2:3])
```

Plotting temperature in Farenheit for Colorado on January 1st.
```{r plot}
# convert to farenheit
temp_f <- (temp_k_regrid *(9/5) - 459.67)

# plot 
ggplot(data= grid_coords, aes(x=lon, y=lat, color = temp_f)) +
  geom_point() +
  ggtitle("Temperature (F) on Jan 1 2015 for Colorado") +
  theme_minimal()
```

Looks good. I'm going to itterate through each temperature nc file to regrid to the Colorado grid. However, I did some quick benchmarks and this script would take nearly 42 minutes to run for one year because of the rasterize function. I'm going to write this in a Rscript to run in parallel on the ozone server. I am going to save the krig grid coordinates for Colorado and run this script from the meta_wildfire repository though to make this easier. 
```{r save_coords}
write_csv(grid_coords, "../../data/smoke/colo_krig_grid_coords.csv")
```

