---
title: "Regrid WRF Grid ID to US 15x15 km Grid"
author: "Ryan Gan"
date: "2018-03-23"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Introduction

I have a 3x3 km resolution grid I think for the entire state of Colorado, where I also have an ID that can joined to health outcomes data. I have exposure estimates at a larger grid size of 15x15 km. I think the most straight-forward approach for this would be to convert each shapefiles to gridded rasters, then extract the ID values of the larger GRID IDs to the smaller WRFGRID points. I could then make a dataframe of WRGRID IDs and their corresponding GRID IDs.

Load "sf" and "raster" spatial library.
```{r library}
library(tidyverse)
library(sf)
library(raster)
```

```{r read_wrfgrid_shape}
wrfgrid_sf <- st_read(dsn = "../../data/shapefiles/wrf_grid", 
                      layer = "wrf_grid")
```

```{r read_grid_shape}
us_grid_sf <- st_read(dsn = "../../data/shapefiles/1016-us_pm_grid",
                      layer = "1016-us_pm_grid")
```

Extracting WRFGRID centroid latitude and longitude coordinates.

```{r wrf_coords}
# extract wrfgrid id vector
wrfgrid_id <- wrfgrid_sf$WRFGRID_ID
# extract centroid points from wrfgrid sf
coords <- do.call(rbind, st_geometry(st_centroid(wrfgrid_sf)))
colnames(coords) <- c("lon", "lat")
# bind both together in spatial dataframe
wrf_coords <- data.frame(cbind(wrfgrid_id, coords))
head(wrf_coords)
```

Subsetting US grid to grid cells in Colorado that overlap the WRFGRID shapefile.

```{r co_grid_subset}
# subset to colorado
co_grid <- us_grid_sf[wrfgrid_sf, ]
```

Plotting Colorado grid and WRFGRID coordinate points.

Extracting Kate's krig grid centroid latitude and longitude coordinates.

```{r krig_coords}
# extract wrfgrid id vector
GRID_ID <- co_grid$GRID_ID
# extract centroid points from wrfgrid sf
coords <- do.call(rbind, st_geometry(st_centroid(co_grid)))
colnames(coords) <- c("lon", "lat")
# bind both together in spatial dataframe
krig_coords <- data.frame(cbind(GRID_ID, coords))
head(krig_coords)
# write this as a csv to use
write_csv(krig_coords, '../../data/smoke/krig_grid_id_coords.csv')
```

```{r plot}
# plot
ggplot(co_grid) +
  geom_sf(fill = NA) +
  geom_point(data=wrf_coords, aes(x=lon, y=lat), 
             color = "blue", size = 0.1, alpha =0.3) +
  theme_minimal()
```

Joining polygons to grid. First step is to create a sf object of the st_centroid. This is similar to the step above, but I retain the geometry for each WRFGRID ID.

```{r wrf_coord_sf}
wrf_coord_sf <- st_centroid(wrfgrid_sf) %>% 
  dplyr::select(WRFGRID_ID)
```

I can now use the st_join function to join the GRID IDS of the Colorado grid with the WRFGRID IDs.
```{r st_join}
grid_id_key <- st_join(co_grid, wrf_coord_sf) %>% 
  filter(!is.na(WRFGRID_ID))
```

There are some missing WRFGRID_IDs for some GRIDS, but those grids look like edge grids.

I'm going to read in Colorado county shapefiles and plot the grids that intersect Denver county. subset the first 10 WRFGRID IDs and plot.

```{r read_denver_county}
# read denver county
denver_sf <- st_read(dsn = "../../data/shapefiles/colorado_county",
                     layer = "colorado_county") %>% 
  filter(NAME == "Denver")
```

Finding the grids that interesect Denver county.

```{r subset_check}
# assign crs from denver county to the grid sf
st_crs(grid_id_key) <- st_crs(denver_sf)
# subset grids that touch denver county
grid_check <- grid_id_key[denver_sf,]
# wrf id check
wrf_check <- wrfgrid_sf %>% 
  filter(WRFGRID_ID %in% grid_check$WRFGRID_ID)
# assign crs
st_crs(wrf_check) <- st_crs(denver_sf)
# wrf_coords points
wrf_check_coords <- wrf_coord_sf %>% 
  filter(wrfgrid_id %in% wrf_check$WRFGRID_ID)
# assign crs
st_crs(wrf_check_coords) <- st_crs(denver_sf)
```

Plotting the 15 km^2^ grid and 3 km^2^ WRF grid and centroids over Denver County.

```{r plot_check}
ggplot(denver_sf) +
  geom_sf(fill = NA, color = "black") +
  geom_sf(data = grid_check, fill = NA, color = "blue") +
  geom_sf(data = wrf_check, fill = NA, color = "red") +
  geom_sf(data = wrf_check_coords, fill = NA, color = "red", size = 0.1) +  
  theme_minimal()
```

Looks like this worked. I'm going to save the key with WRFGRID coordinates to make sure I assign grid PM~2.5~ values correctly to WRFGRID IDs.

```{r save_wrf_ids, eval=F}
# join with coords
gridid_key_coords <- grid_id_key %>% 
  left_join(wrf_coords, by = c("WRFGRID_ID" = "wrfgrid_id")) %>% 
  dplyr::select(GRID_ID, WRFGRID_ID, lon, lat, COLX, ROWY)
# remove geometry
st_geometry(gridid_key_coords) <- NULL
# write csv
write_csv(gridid_key_coords, path = "../../data/shapefiles/wrfgrid_key.csv")
```

