---
title: "WRF grid to FRAPPE Grid"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

R notebook that reduces Colorado WRF grid size to FRAPPE extent.

```{r load_library}
library(tidyverse)
library(sf)
```

Read in Colorado WRF grid and the coordinates Wande sent me.

```{r read_files}
# wrf grid
wrfgrid <- st_read(dsn = "../../data/shapefiles/wrf_grid/",
                   layer = "wrf_grid")
# coordinates
frappe_coord <- read_delim("../../data/shapefiles/CO_3k_latlong.txt", 
                           delim = " ")
```

Bounding box of WRF grid shapefile.
```{r wrf_bbox}
st_bbox(wrfgrid)
```

Extent of Wande's coordinates.

```{r coord_extent}
# min
c(min(frappe_coord[,1]), min(frappe_coord[,2]))
# max
c(max(frappe_coord[,1]), max(frappe_coord[,2]))
```

I want to retain the WRFGRID IDs as well as the coordinates. Using custom clip function of simple feature.

```{r clip_function}
bbox_clip <- function(sf, bbox) {
  # find the CRS of the sf object
  crs <- sf::st_crs(sf)$proj4string
  # create matrix
  x <- c(bbox[1], bbox[1], bbox[3], bbox[3], bbox[1])
  y <- c(bbox[2], bbox[4], bbox[4], bbox[2], bbox[2])
  coords <- matrix(cbind(x, y), ncol=2)
  # create polygon and assign same coord crs as sf object
  coords_poly <- sp::Polygon(coords)
  bbox_poly <- sp:: SpatialPolygons(list(sp::Polygons(list(coords_poly),
    ID = "bbox")), proj4string = sp::CRS(crs))
  # convert to sf feature
  bbox_sf <- st_as_sf(bbox_poly)
  # clip sf object
  clipped_sf <- sf[bbox_sf,]
  return(clipped_sf)
}
```

Define new bounding box based on coordinates Emily provided.

```{r new_bbox}
# new bbox
frappe_bbox <- c(-106, 38.3687, -101.3326, 41.22053)
# new grid
new_grid <- bbox_clip(wrfgrid, frappe_bbox)
```

Build coordinate matrix with WRFGRID ID. 

```{r extract_coords}
# convert geometry of new coords to a list
geo_list <- st_geometry(new_grid) %>% as.list()
# extract nw coords from each grid
nw_coords <- lapply(geo_list, function(x){
        nw_lon = unlist(st_geometry(x))[2]
        nw_lat = unlist(st_geometry(x))[7]
        data = data.frame(nw_lon, nw_lat)
  }) %>% 
  do.call(rbind, .)

# wrfgrid id vector
new_wrfid <- new_grid$WRFGRID_ID
# frapp grid
frappe_grid <- data.frame(new_wrfid, nw_coords) %>% 
  rename(WRFGRID_ID = new_wrfid)

# check new grid
ggplot(frappe_grid, aes(x=nw_lon,y=nw_lat)) +
  geom_point(color = "red", size = 0.1) +
  geom_point(data=frappe_coord, aes(x=long,y=lat), color = "blue", size = 0.1,
             alpha = 0.2) +
  theme_minimal()
```

Write file for Emily.

```{r frappe_grid_csv}
write_csv(frappe_grid, "../../data/shapefiles/frappe_grid.csv")
```

