---
title: "Creating Colorado daily grid PM~2.5~ and temperature 2010-2015"
author: "Ryan Gan"
date: '2018-03-22'
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: inline
---

## Introduction

This R markdown file combines daily estimates of the kriged PM~2.5~, seasonal background PM~2.5~, HMS, and temperature from each year and creates a time series data frame by GRID ID. I am considering putting this in a SQLite database along with zip-level and county-level values to make it easier to join with outcomes.

```{r library}
library(tidyverse)
```

## PM~2.5~ 

Defining file path where files for PM~2.5~, background, and HMS data are stored.

```{r pm_path}
# file path
pm_file_path <- paste0("../../data/smoke/1016-co_krig_grid_pm/")
```

Importing PM~2.5~ estimated using ordinary kriging. I'm going to need to provide a different name to distinguish these values from county and zip values. I'm going call it pm25_grid.

```{r krig_pm25}
# list files
pm25_grid <- list.files(pm_file_path, pattern = "krig_pm") %>% 
  # read in each csv file using map
  map( ~ read_csv(paste0(pm_file_path, .),
                  col_types = cols(.default = "d", GRID_ID = "c"))) %>% 
  # transform data wide to long
  map_dfr( ~ gather(., date, pm_krig, -GRID_ID) %>% 
             mutate(date = as.Date(str_sub(date, start=2), format = "%Y%m%d"),
                    pm25_grid = as.numeric(pm_krig)) %>% 
             select(-pm_krig))
# glimpse rows
glimpse(pm25_grid)
```

## Seasonal Background PM~2.5~

Importing seasonal background PM~2.5~. I'm going to label this seasonal back ground PM~2.5~ for the grid as sbg_pm_grid.

```{r backround_pm25}
# list files
bg_pm25_grid <- list.files(pm_file_path, pattern = "background_pm") %>% 
  # read in each csv file using map
  map( ~ read_csv(paste0(pm_file_path, .),
                  col_types = cols(.default = "d", GRID_ID = "c"))) %>% 
  # transform data wide to long
  map_dfr( ~ gather(., date, bg_pm, -GRID_ID) %>% 
             mutate(date = as.Date(str_sub(date, start=2), format = "%Y%m%d")) %>% 
             rename(sbg_pm_grid = bg_pm))
# glimpse rows
glimpse(bg_pm25_grid)
```

## HMS Smoke

Hazard mapping system smoke in atmospheric column. Binary variable where 1 is yes and 0 is no. I'm going to lable this as hms_grid.

```{r hms}
# list files
hms_grid <- list.files(pm_file_path, pattern = "hms") %>% 
  # read in each csv file using map
  map( ~ read_csv(paste0(pm_file_path, .),
                  col_types = cols(.default = "d", GRID_ID = "c"))) %>% 
  # transform data wide to long
  map_dfr( ~ gather(., date, hms, -GRID_ID) %>% 
             mutate(date = as.Date(str_sub(date, start=2), format = "%Y%m%d")) %>% 
             rename(hms_grid = hms))
# glimpse rows
glimpse(hms_grid)
```

## Temperature

Reading in temperature in kelvin and converting to farenheit.

```{r temp}
# temp path
temp_path <- paste0("../../data/smoke/co_air_temp/")
# list files
temp_grid <- list.files(temp_path) %>% 
  # read in each csv file using map
  map( ~ read_csv(paste0(temp_path, .),
                  col_types = cols(.default = "d", GRID_ID = "c"))) %>% 
  # transform data wide to long
  map_dfr( ~ gather(., date, temp_k, -GRID_ID) %>% 
             mutate(date = as.Date(str_sub(date, start=2), format = "%Y%m%d"),
                    temp_f_grid = round(temp_k * (9/5) - 459.67, 2)) %>% 
             select(-temp_k))
# glimpse rows
glimpse(temp_grid)
```

## Composite Data

Binding all data frames together.

```{r comp_data}
# create list of all dataframes
co_grid_env_ts <- list(pm25_grid, bg_pm25_grid, hms_grid, temp_grid) %>% 
  # bind list together
  reduce(function(df1, df2) left_join(df1, df2, by = c("GRID_ID", "date")))
# glimpse df
glimpse(co_grid_env_ts)
```

Writting file. Option of eval = F so I don't keep running it if I knit.

```{r grid_write, eval = F}
write_csv(co_grid_env_ts, "../../data/smoke/1015-grid_pm.csv")
```

