---
title: "Colorado claims"
author: "Jingyang Liu"
date: "February 3, 2017"
output: html_document
---


## Overview
This document contains results for the association between wildfire smoke PM~2.5~ and cardiovascular (CVD) and respiratory health outcomes. This particular markdown document focuses on the zip code-level analysis. Note there is a county-level analysis that will be paired with CDC mortality data as well. Herein, I compare different PM~2.5~ estimation methods described in further detail below. 

Note, epidemiological methods that Sheryl and Ryan think can improve our estimation and reduce some bias around the point estimates will not be the focus of the manuscript. 

```{r setup, include=FALSE}



# ------------------------------------------------------------------------------
# Title: Script descriptive statistics about the Colorado claims data more than just 2012. 
#  some counts of the primary diagnoses on our specific health outcomes over that time frame.
# Author: Jingyang Liu
# Date Created: 1/24/17
# R version: 3.3.2
# ------------------------------------------------------------------------------


# loading libraries used
library(tidyverse) # import tidyverse
library(survival) # for conditional logistic regression
library(htmlTable) # table
library(lme4) # random-effects model
library(broom) # broom for tidy data from stats models
library(cowplot) # for multi plot image



# Set working directory and read in files --------------------------------------
# relative path

path <- paste0("./data/case_cross_data/") # for some reason, I cannot use this command 
# for the following use, so I have to write the working directory path 12 times.

# direct path for pc
#path <- paste0("C:/Users/RGan/Documents/git_local_repos/wildfire/",
#               "wildfire_washington/analysis/analysis_data")

# Infile case-crossover dataframes ---------------------------------------------
# Dataframes made in 'chars_3month_binary_smoke_may2016 script
# resp exacerbations
resp_casecross <- read_csv(paste(path, 
                                 "resp1_jul_to_oct_casecross.csv", 
                                 sep = "/"))

# asthma
asthma_casecross <- read_csv(paste(paste0("./data/case_cross_data/"), 
                                   "asthma1_jul_to_oct_casecross.csv", 
                                   sep = "/")) 

getwd()
setwd("./data/cas_cross_data")
# copd 
copd_casecross <- read_csv(paste(paste0("./data/case_cross_data/"), 
                                 "copd1_jul_to_oct_casecross.csv", 
                                 sep = "/"))
# copd exacerbations
#copd_ex_casecross <- read_csv(paste(paste0("./data/case_cross_data/"), 
#                                    "copd_ex1_jul_to_oct_casecross.csv",
#                                    sep="/"))
# pneum or bronchitis
pneum_casecross <- read_csv(paste(paste0("./data/case_cross_data/"), 
                                  "pneum1_jul_to_oct_casecross.csv",
                                  sep="/"))
# acute bronchitis
acute_bronch_casecross <- read_csv(paste(paste0("./data/case_cross_data/"), 
                                         "acute_bronch1_jul_to_oct_casecross.csv",
                                         sep = "/"))
# cvd
cvd_casecross <- read_csv(paste(paste0("./data/case_cross_data/"), 
                                "cvd1_jul_to_oct_casecross.csv",
                                sep="/"))
# arrhythmia
arrhythmia_casecross <- read_csv(paste(paste0("./data/case_cross_data/"), 
                                       "arrhythmia1_jul_to_oct_casecross.csv",
                                       sep="/"))
# cerebral vascular
cereb_vas_casecross <- read_csv(paste(paste0("./data/case_cross_data/"), 
                                      "cereb_vas1_jul_to_oct_casecross.csv",
                                      sep="/"))
# heart failure
hf_casecross <- read_csv(paste(paste0("./data/case_cross_data/"), 
                               "hf1_jul_to_oct_casecross.csv", 
                               sep="/"))
# ischemic heart disease
ihd_casecross <- read_csv(paste(paste0("./data/case_cross_data/"), 
                                "ihd1_jul_to_oct_casecross.csv",
                                sep="/"))
# myo infarc
mi_casecross <- read_csv(paste(paste0("./data/case_cross_data/"), 
                               "mi1_jul_to_oct_casecross.csv", sep="/"))

# broken arm
broken_arm_casecross <- read_csv(paste(paste0("./data/case_cross_data/"), 
                                       "broken_arm1_jul_to_oct_casecross.csv",
                                       sep="/"))


# note 12/4/16: read_csv is pretty slow on pc, fread doesn't import date variable 
# that well. Solution may be to use fread but convert to date to use later
```


## Descriptive

```{r descriptive table, echo = F, warning = F, results='asis'}
# dataframe list
df_list <- list(resp_casecross, asthma_casecross, copd_casecross, pneum_casecross,
                acute_bronch_casecross, cvd_casecross, arrhythmia_casecross,
                cereb_vas_casecross, hf_casecross, ihd_casecross, mi_casecross, 
                broken_arm_casecross)

outcome_list <- c('All Respiratory', 'Asthma', 'COPD', 'Pneumonia', 
                  'Acute Bronchitis', 'Cardiovascular Disease', 'Arrhythmia', 
                  'Cerebrovascular Disease', 'Heart Failure',
                  'Ischemic Heart Disease', 'Myocardial Infarction', 'Broken Arm')

# create an empty list to row bind dataframes together
datalist <- list()

# set list and do cross tabs to find values
# Producing conditional logit model estimates loop 
for(i in 1:length(df_list)){
  
  # dataframe to loop through
  df_to_loop <- data.frame(df_list[i])
  # indication of column
  outcome <- colnames(df_to_loop[13])
  # outcome name
  outcome_name <- outcome_list[i]
  
  # dataframe for analysis creation
  
  df_analysis <- df_to_loop %>% 
    # only look at outcomes
    filter(outcome == 1) %>%
    # add another row that makes sure there is a person <15 in the dataframe
    # tricking xtabs to produce a 0 cell for the outcome for age <15
    add_row(outcome = 0, age_ind = 0)
  
  # cross tabs
  outcome_n <- xtabs(~ outcome, df_analysis)
  cross_tab_age <- xtabs(~ outcome + age_ind, df_analysis)
  cross_tab_sex <- xtabs(~ outcome + sex_ind, df_analysis)
  # empty matrix
  point_estimates <- matrix(nrow = 1, ncol = 7, byrow = T)
  
  colnames(point_estimates) <- c("outcome", "n", "age_15", "age_15_65", 
                                 "age_65", "female", "male")
  
  # fill in the outcome name for the dataframe before the loop
  point_estimates[, 1] <- outcome_name
  # fill n
  point_estimates[, 2] <- outcome_n[2] # second element of the 1 dimension vector
  # age <15
  point_estimates[, 3] <- cross_tab_age[2, 1]
  # age 15 to 65
  point_estimates[, 4] <- cross_tab_age[2, 2]
  # age >65
  point_estimates[, 5] <- cross_tab_age[2, 3]
  # male == 0
  point_estimates[, 7] <- cross_tab_sex[1, 1]
  # female == 1
  point_estimates[, 6] <- cross_tab_sex[1, 2]
  
  
  # save point estimates as a dataframe
  point_est_df <- as_data_frame(point_estimates)
  
  # combine previous values in dataframe that has all outcome/methods comparisons
  datalist[[i]] <- point_est_df
  
} # end of loop

# combine each outcome dataframe itteration in to a big dataset
combined_point_est_df <- bind_rows(datalist) %>% 
  # find proportions/percents for each strata in a row
  mutate(age_15_pr = as.character(round((as.numeric(age_15)/as.numeric(n))*100,1)),
         age_15_65_pr = as.character(round((as.numeric(age_15_65)/as.numeric(n))*100,1)),
         age_65_pr = as.character(round((as.numeric(age_65)/as.numeric(n))*100,1)),
         female_pr = as.character(round((as.numeric(female)/as.numeric(n))*100,1)),
         male_pr = as.character(round((as.numeric(male)/as.numeric(n))*100,1))) %>% 
  select(outcome, n, age_15, age_15_pr, age_15_65, age_15_65_pr, age_65,
         age_65_pr, female, female_pr, male, male_pr)

#str(combined_point_est_df)

tab <- htmlTable(txtRound(combined_point_est_df, digits = 1), 
                 caption = "Number of cases for each outcome observed from July 1st to October 31st, 2012",
                 # column headers
                 header = c("Outcome", "Cases n", "Less than 15", "(%)", "15 to 65", "(%)", "Greater than 65",
                            "(%)", "Female", "(%)", "Male", "(%)"),
                 # column spanner
                 cgroup = c("","Age Category", "Sex"), 
                 n.cgroup = c(2, 6, 4),
                 padding.rgroup = "&nbsp;&nbsp;",
                 css.cell = "padding-left: 0.5em; padding-right: .5em;", # cell space
                 align = "llccccc" # column alignment,
) # end table


print(tab)

```


