---
title: "ZIP code emergency department rates"
author: "Ryan Gan"
date: "11/1/2017"
output: html_document
---
## Document background/overview

As part of the Colorado chapter of the American Lung Assocition's project to qunatify the health impacts of two coal-fired power plants in Colorado Springs and Pueblo, we need to calculate baseline cardiopulmonary emergency department (ED) visit rates in Colorado from 2010 to September of 2015. We've decided to limit the period of estimates from 2010 to 2014. I also had issues with calculating at the census tract level so I'm going to try ZIP code.

Set up. Calling tidyverse and sf packages.
```{r setup, echo=F, message=F, warning=F}
library(tidyverse)
library(sf)
```

Importing county denominators and simple features object as well as hospital claims data. 

```{r data import, echo=F, message=F, warning=F}
# read in the dataframe with outcomes 
hosp_path <- "../data/health/co_hosp_w_outcome_df.csv"
co_hosp <- read_csv(hosp_path) %>% 
  # noticed broomfield county_geo 014 was sometimes assigned county_final 159
  # there is no county 159. All other spatial indicators suggest broomfield 014
  mutate(county_final = ifelse(county_geo == "014" & county_final == "159", 
                               "014", county_final),
    # make fips code
    FIPS = paste0("08", county_final),
    # adding in a season variable for estimation by season 
    season = as.factor(ifelse(ADMMM >= 3 & ADMMM <= 5, "spring",
             ifelse(ADMMM >= 6 & ADMMM <= 8, "summer",
               ifelse(ADMMM >= 9 & ADMMM <= 11, "fall", "winter")))),
  # adding in a year indicator for season to check for trends
  season_yr = as.factor(paste0(season, ADMYY)),
    # set wrfgrid values of 0 to missing
    WRFGRID_ID = ifelse(WRFGRID_ID == 0, NA, WRFGRID_ID),
    WRFGRID_ID = as.character(WRFGRID_ID)) %>% 
  # remove those missing a WRFGRID ID
  filter(!is.na(WRFGRID_ID))

# import census denominators for counties and states ----
census_estimates <- read_csv("../data/health/co-est2016-alldata.csv") %>% 
  # make fips variable
  mutate(FIPS = paste0(STATE, COUNTY))

# lat/lon of power plants ----
plant_id <- c("pueblo", "colo_springs")
lat <- c(38.2081, 38.8244)
lon <- c(-104.5747, -104.8331)

plant_locations <- tibble(plant_id, lat, lon) 
```

Read in ZIP and county simple features that will be used for maps latter.
```{r simple features import, echo=F, warning=F, message=F}
# read zip shapefile Sheena sent 
co_zip_2014_path <- "../data/shapefiles/2014_ZCTA_Population"
co_zip_2014_sf <- st_read(dsn=co_zip_2014_path, layer="CO_ZCTA_2014", quiet = T)
# save crs 
nad83 <- st_crs(co_zip_2014_sf)

# read in populations
zip_pop <- read_csv(paste0("../data/shapefiles/2014_ZCTA_Population/",
  "Populations_CO_ZCTA_2014.txt")) %>% 
  select(ZCTA5CE10, B01001e1) %>% 
  # need to only extract the last part of the GEOID to join 
  rename(ZIP = ZCTA5CE10,
    total_pop = B01001e1) %>% 
  mutate(ZIP = as.character(ZIP))

# read county shapefile
county_path <- "../data/shapefiles/colorado_county"
county_sf <- st_read(dsn=county_path, layer = "colorado_county") %>% 
  # make fips variable
  mutate(FIPS = paste0(STATEFP, COUNTYFP)) %>% 
  # assign nad83
  st_transform(nad83)

# subset to study area 
# pikes peak study area: Teller and El Paso
# soutern central: Custer, Huerfano, Pueblo, and Las Animas
ala_study_sf <- county_sf %>% 
  filter(NAME %in% c("Teller", "El Paso", "Custer", "Huerfano", 
                   "Pueblo", "Las Animas")) %>% 
  st_transform(nad83)
```

## ZIP rate estimates 2010-2014

Estimating 2010-2014 period rates for census tracts in Colorado and then subsetting to the ALA study area to make viewing easier.

```{r zip period rate, echo=F, message=F, warning=F}
zip_rate_period <- co_hosp %>% 
  filter(ADMYY != 15) %>% 
  group_by(ZIP) %>% 
  summarise(cvd_n = sum(cvd_dx), resp_n = sum(resp_dx)) %>% 
  # filter to zip codes in colorado
  filter(as.numeric(ZIP) >= 80001 & as.numeric(ZIP) <= 81658) %>% 
  # join with census denom
  right_join(zip_pop, by = "ZIP") %>% 
  # removing zip codes with 0 population
  filter(total_pop != 0) %>% 
  # calculate rate 
  mutate(cvd_n = ifelse(is.na(cvd_n), 0, cvd_n),
  resp_n = ifelse(is.na(resp_n), 0, resp_n),
  # set NA cvd resp n to 0
  cardiopulm_n = cvd_n + resp_n,
  total_pop5y = total_pop*5,
  cardiopulm_per_100p5y = (cardiopulm_n/total_pop5y)*100,
  cvd_per_100p5y = (cvd_n/total_pop5y)*100,
  resp_per_100p5y = (resp_n/total_pop5y)*100)

# join with zip sf
zip_rate_sf <- co_zip_2014_sf %>% 
  rename(ZIP = ZCTA5CE10) %>%
  mutate(ZIP = as.character(ZIP)) %>% 
  left_join(zip_rate_period, by = "ZIP")
```

### ZIP code CVD rate map 2010-2014

```{r ala cvd rate map, echo=F, message=F, warning=F}
# ALA study area
ala_sf <- st_intersection(zip_rate_sf, ala_study_sf)

# plot
p <- ggplot(ala_sf) +
  geom_sf(aes(fill=cvd_per_100p5y), colour = "white") +
  scale_fill_gradient(name="CVD Rate", low = "#cbb4d4", high = "#20002c") +
  geom_sf(data=ala_study_sf, alpha = 0) +
  geom_point(data=plant_locations, aes(x=lon,y=lat), colour= "red") +
  ggtitle("ALA ZIP CVD ED rates per 100 persons from 2010 to 2014") +
  theme_bw()
# print map
p
```

### ZIP code respiratory rate map

```{r ala resp rate map, echo=F, message=F, warning=F}
# plot
p <- ggplot(ala_sf) +
  geom_sf(aes(fill=resp_per_100p5y), colour = "white") +
  scale_fill_gradient(name= "Resp Rate", low = "#26d0ce", high = "#1A2980") +
  geom_sf(data=ala_study_sf, alpha = 0) +
  geom_point(data=plant_locations, aes(x=lon,y=lat), colour= "red") +
  ggtitle("ALA ZIP respiratory ED rates per 100 persons from 2010 to 2014") +
  theme_bw()
# print map
p
```

### El Paso CVD ED rate 2010-2014

Limiting to El Paso to view Colorado Springs area.

```{r el paso cvd map, echo=F, message=F, warning=F}
el_paso_sf <- ala_study_sf %>% 
  filter(NAME == "El Paso") %>% 
  st_transform(nad83)

el_paso_zip_sf <- st_intersection(zip_rate_sf, el_paso_sf)

# cs plant
cs_plant <- plant_locations %>% filter(plant_id == "colo_springs")

# plot
p <- ggplot(el_paso_zip_sf) +
  geom_sf(aes(fill=cvd_per_100p5y), colour = "white") +
  scale_fill_gradient(name = "CVD Rate", low = "#cbb4d4", high = "#20002c") +
  geom_sf(data=el_paso_sf, alpha = 0) +
  geom_point(data=cs_plant, aes(x=lon,y=lat), colour= "red") +
  ggtitle("El Paso CVD ED rates per 100 persons from 2010 to 2014") +
  theme_bw()
# print map
p
```

### El Paso respiratory ED rate 2010-2014

```{r el paso resp map, echo=F, message=F, warning=F}
# plot
p <- ggplot(el_paso_zip_sf) +
  geom_sf(aes(fill=resp_per_100p5y), colour = "white") +
  scale_fill_gradient(name = "Resp Rate", low = "#26d0ce", high = "#1A2980") +
  geom_sf(data=el_paso_sf, alpha = 0) +
  geom_point(data=cs_plant, aes(x=lon,y=lat), colour= "red") +
  ggtitle("El Paso Resp ED rates per 100 persons from 2010 to 2014") +
  theme_bw()
# print map
p
```

```{r ed rate season, warning=F, message=F, echo=F}
zip_rate_season <- co_hosp %>% 
  filter(ADMYY != 15) %>% 
  group_by(ZIP, season) %>% 
  summarise(cvd_n = sum(cvd_dx), resp_n = sum(resp_dx)) %>% 
  # filter to zip codes in colorado
  filter(as.numeric(ZIP) >= 80001 & as.numeric(ZIP) <= 81658) %>% 
  # join with census denom
  right_join(zip_pop, by = "ZIP") %>% 
  # removing zip codes with 0 population
  filter(total_pop != 0) %>% 
  # calculate rate 
  mutate(cvd_n = ifelse(is.na(cvd_n), 0, cvd_n),
  resp_n = ifelse(is.na(resp_n), 0, resp_n),
  # set NA cvd resp n to 0
  cardiopulm_n = cvd_n + resp_n,
  total_pop15m = total_pop*1.25,
  cardiopulm_per_100p15m = (cardiopulm_n/total_pop15m)*100,
  cvd_per_100p15m = (cvd_n/total_pop15m)*100,
  resp_per_100p15m = (resp_n/total_pop15m)*100)
```

### Summary of max and median rate per 100 persons for seasonal period (1.25 years or 15 months)

Subsetted to ALA study region. Minimum rate for both CVD and respriatory ED rates is 0. May have sparse data problem at seasonl level.  

```{r season summary, echo=F, warning=F, message=F}
season_summary <- zip_rate_season %>% 
  # filter to ala study area
  filter(ZIP %in% ala_sf$ZIP) %>% 
  group_by(season) %>% 
  summarise(max_cvd_rate = max(cvd_per_100p15m), 
    med_cvd_rate = median(cvd_per_100p15m),
    max_resp_rate = max(resp_per_100p15m), 
    med_resp_rate = median(resp_per_100p15m)) %>% 
  filter(!is.na(season)) 

# print rate
season_summary
```