---
title: "Extracting Grid PM~2.5~ Values for Colorado"
author: "Ryan Gan"
date: "2018-03-23"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

## Introduction

I have a 3x3 km resolution grid I think for the entire state of Colorado, where I also have an ID that can joined to health outcomes data. I have exposure estimates at a larger grid size of 15x15 km. I think the most straight-forward approach for this would be to convert each shapefiles to gridded rasters, then extract the ID values of the larger GRID IDs to the smaller WRFGRID points. I created a csv file that contains the WRFGRID IDs in the larger GRID IDs that I will use to assign extracted grid values from the netcdf files Kate provided to WRFGRID IDs.

I think I need tidyverse and ncdf4.
```{r libraries}
library(tidyverse)
library(ncdf4)
library(sf)
```

### NetCDF PM~2.5~ Values

Before I read in PM~2.5~ values from the netcdf files, I'm going to load the WRFGRID to GRID key I created in the script "regrid_wrf.Rmd". I read all variables as characters and then converted to numeric as I was getting some parsing warning messages that I think are fine.

```{r grid_key}
# read values
grid_key <- read_csv("../../data/shapefiles/wrfgrid_key.csv", 
                     col_types = cols(.default = "c")) %>% 
  mutate_all(as.numeric)
# viewing first lines of tibble
head(grid_key)
```

I'm going to start by defining path to nc files in the meta_wildfire repo and listing the files. I have PM~2.5~ values for every year for the entire US from 2006 to 2015. I only need the grid values for Colorado from 2010 to 2015, as those are the years I have data. 

```{r nc_path}
# define file path
nc_file_path <- paste0("../../../meta_wildfire/data/smoke/krigedPM25_06-15/")
# create vector of list file names
nc_names <- list.files(nc_file_path)
# printing names in vector
nc_names
```

I'm going to start by testing out how to extract only values I need for 2010. Opening connection to nc file and printing nc summary.

```{r open_nc_connection}
# open nc connection to 2010 nc file
pm_2010_nc <- nc_open(paste0(nc_file_path, "krigedPM25_2010.nc"))
# print nc summary
pm_2010_nc
```

It looks like Kate saved the row and columns that I need to output as variables. I'm going to see if I can use these.

Range of column/x cells. I need to add 1 to each value to get the right location in the array since the values start at 0.
```{r output_x}
# output unique cell locations for column/x
colx <- unique(grid_key$COLX) + 1
# range of column x
range(colx)
```

Range of row/y cells. I need to add 1 to each value for hte same reason above.
```{r output_y}
# output unique cell lcoations for row/y
rowy <- unique(grid_key$ROWY) + 1
# range of row y
range(rowy)
```

### Checking Point PM~2.5~ 

Attempting to output the PM~2.5~ values for January 1st, 2010 for the specific grid cells over Colorado. 

```{r pm_jan1}
# colorado pm values
pm <- ncvar_get(pm_2010_nc, varid = "PM25")[104:147, 81:112, 1]
# lat locations
lat <- ncvar_get(pm_2010_nc, varid = "lat")[104:147, 81:112]
# lon locations
lon <- ncvar_get(pm_2010_nc, varid = "lon")[104:147, 81:112]
# bind vectors together
pm_jan1_df <- data.frame(as.vector(pm), as.vector(lat), as.vector(lon))
colnames(pm_jan1_df) <- c("pm", "lat", "lon")
```

Plotting PM~2.5~ values over Colorado shapefile from maps package. Projection may be slightly off since the shapefile likely has a WGS84 or NAD83 projection and the lat/lon coordinates of the PM~2.5~ values doe not.

```{r colorado_plot}
# read colorado shape
colorado <- map_data("state", region = "colorado")

# plot
plot <- ggplot(data = pm_jan1_df, aes(x = lon, y = lat, color = pm)) +
  geom_point() +
  geom_polygon(data=colorado, aes(x = long, y = lat, group = group), fill = NA, 
               color = "red") + 
  ggtitle("Grid location PM2.5 values on Janurary 1st, 2010 for Colorado") +
  theme_minimal()

plot
```

I can't join by latitude/longitude, so I'm seeing if I can expand the colx and rowy locations to match the locations provided in the shapefile by Bonne.

### Checking PM~2.5~ Assignment to GRID ID

I believe the GRID_IDs should start in the lower-left and go left to right, bottom to top. Joining numbered sequences to the PM coordinates and plotting locations of text to see. 

```{r point_plot}
# finding out where cell ids are 
# pretty sure they will start in the lower left corner going right, then up rows
pm_jan1_df$cell_id <- 1:1408
# plot
ggplot(pm_jan1_df, aes(x=lon, y=lat)) +
  geom_text(aes(label = cell_id), size = 1.5) +
  theme_minimal()
```

It goes left to right, bottom to top. So I will need to make sure the vector of GRID_IDs matches the order of the vector of daily PM values.

```{r new_colorado_shape}
# define grid path
grid_path <- paste0("../../data/shapefiles/co_krig_grid/")

# read western US grid and subset to the COLX and ROWY values in xy expanded grid
co_grid_sf <- st_read(dsn = grid_path, layer = "co_krig_grid") %>% 
  arrange(ROWY, COLX)
```

Sorting column x and row y to make sure it matches the vector of cell numbers of the Colorado grid values. Extract vector of the 1408 GRID IDs to assign PM~2.5~ to. Plotting first 50 GRID_IDs of the arragnged shapefile to make sure it looks correct.

```{r grid_plot}
# checking grid
grid_check <- st_centroid(co_grid_sf) %>% slice(1:50)
# plot
ggplot(grid_check) +
  geom_sf() +
  xlim(-110, -101) +
  ylim(36, 41.2) +
  theme_minimal()
```

This arranged/sorted simple features shapefile should be correct for when I extract PM~2.5~ values.

Before moving on, I want to check that the values for the assigned grid ids look correct.
```{r checking_jan1_sf}
# attempt to extract daily estimates for january for colorado from nc file
pm_jan1_vec <- as.vector(ncvar_get(pm_2010_nc, varid = "PM25")[104:147, 81:112, 1])
# check to see if equal with pm_jan1_df
identical(pm_jan1_vec, pm_jan1_df$pm)
```

Joining grid id, creating a dataframe, joining to the Colorado grid simple features, and plotting.

```{r sf_plot_check}
# output sorted grid id
grid_id <- co_grid_sf$GRID_ID
# create dataframe 2
pm_jan1_sf <- data.frame(grid_id, pm_jan1_vec) %>% 
  rename(GRID_ID = grid_id, pm = pm_jan1_vec) %>% 
  left_join(co_grid_sf, by = "GRID_ID")
# plot
ggplot(pm_jan1_sf) +
  geom_sf(aes(fill = pm), color = NA) +
  theme_minimal()
```

This plot looks similar to the point plot where there is lower PM~2.5~ in the south-west corner. This looks good. I can extract and create time-series dataframes of values. I think I can run this on a local machine without too much trouble. Giving it a go.

## GRID PM~2.5~ Values 2010 to 2015

I only need values from 2010 to 2015 to match my outcomes data, so I'm going to reduce the list to these years.

```{r pm_list_study}
# reducing names list to 2010 to 2015
nc_list <- nc_names[5:10]
# checking list names
nc_list
```

Itterating to extract PM~2.5~, seasonal background PM~2.5~, and HMS smoke for each grid cell in Colorado and writing as seperate dataframes in a folder.

```{r colorado_grid_pm_dfs, eval=F}
# output sorted grid id to bind in apply function
GRID_ID <- co_grid_sf$GRID_ID

start <- Sys.time()
# apply over vector of nc names
sapply(nc_list, function(meow){
  # open nc connection
  pm_nc <- nc_open(paste0(nc_file_path, meow))
  # output 3d arrays from .nc files for colorado grids and convert to 2d matrices 
  pm_mat <- apply(ncvar_get(pm_nc, varid = "PM25")[104:147, 81:112, ],3, as.vector)
  bg_pm_mat <- apply(ncvar_get(pm_nc, varid = "Background PM25")[104:147, 81:112, ], 3, as.vector)
  hms_mat <- apply(ncvar_get(pm_nc, varid = "HMS Smoke")[104:147, 81:112, ], 3, as.vector)

  # extract year from name of list
  year <- substring(meow, 12, 15)

  # kates date in the .nc file is just one value; i'm just going to create
  # my own date vector that will account for leap year too
  date <- paste0("d", gsub("-", "", 
    seq.Date(from = as.Date(paste0(year, "-01-01")), 
      to = as.Date(paste0(year, "-12-31")), by = "day"))) # end date seq

  # bind in to a list to itterate through population weighting function
  pm_list <- list(pm_mat, bg_pm_mat, hms_mat)
  # assign names to list elements
  names(pm_list) <- c("krig_pm", "background_pm", "hms_smk")
  
  # lapply row and column names and sort by GRID_ID Cell
  pm_list <- lapply(pm_list, function(x){
    # assign date to column name
    colnames(x) <- date
    # assign grid_id to rownames
    rownames(x) <- GRID_ID
    # column-bind GRID_ID to start of matrix
    x <- cbind(GRID_ID, x)
    x <- x[order(as.numeric(row.names(x))),]
    df <- data.frame(x)
  }) # end function applying name

  lapply(1:length(pm_list), function(x){
  # function to write elements in list to folder
      # write_name
    write_name <- paste0("../../data/smoke/1016-co_krig_grid_pm/",
          year, "-co_krig_grid_", names(pm_list[x]),".csv")
    # save csv file
    write_csv(pm_list[[x]], write_name)
  }) # end function writing file

}) # end sapply

stop <- Sys.time()
runtime <- stop - start
runtime
```

Takes about 25 seconds to do this. I set evaluation of this code chunk to false so I don't perform this step again when I knit.
