---
title: "Creating Krig Grid Shapefile for Colorado"
author: "Ryan Gan"
date: "3/26/2018"
output: html_document
---

## Introduction

This document contains code that creates a smaller shapefile of the krig grid cells that overlays the entire contiental United States to just the grids over the state of Colorado and that touch a WRFGRID-ID. This smaller shapefile will make it easier to assign Colorado PM~2.5~ estimats. 

I will need the tidyverse and sf libraries.
```{r libraries}
library(tidyverse)
library(sf)
```

Reading in the grid key that links Krig Grid GRID_ID to the WRF grid WRFGRID_ID 
```{r grid_key}
# read values
grid_key <- read_csv("../../data/shapefiles/wrfgrid_key.csv", 
                     col_types = cols(.default = "c")) %>% 
  mutate_all(as.numeric)
# viewing first lines of tibble
head(grid_key)
```

Range of column/x cells. I need to add 1 to each value to get the right location in the array since the values start at 0.
```{r output_x}
# output unique cell locations for column/x
colx <- unique(grid_key$COLX) + 1
# range of column x
range(colx)
```

Range of row/y cells. I need to add 1 to each value for hte same reason above.
```{r output_y}
# output unique cell lcoations for row/y
rowy <- unique(grid_key$ROWY) + 1
# range of row y
range(rowy)
```

Read in larger US krig grid, subset to Colorado by cell ids in column x and row y.

```{r new_colorado_shape}
# define grid path
grid_path <- paste0("../../../meta_wildfire/data/shapefile/1016-us_pm_grid/")

# read western US grid and subset to the COLX and ROWY values in xy expanded grid
co_grid_sf <- st_read(dsn = grid_path, layer = "1016-us_pm_grid") %>% 
  mutate(col_x = COLX + 1, row_y = ROWY + 1) %>% 
  filter(col_x %in% colx & row_y %in% rowy) %>% 
  dplyr::select(-COLX, -ROWY, -AREA, -PERIMETER) %>% 
  rename(COLX = col_x, ROWY = row_y)
```

Plotting to make sure it looks right.

```{r plot}
ggplot(co_grid_sf) +
  geom_sf(fill =NA)
```

Writing shapefile.

```{r write_sf, eval = F}
# write colorado grid shape
st_write(co_grid_sf, dsn = "../../data/shapefiles/co_krig_grid", 
         layer = "co_krig_grid", driver = "ESRI Shapefile", delete_layer = T)
```