---
title: "R Notebook"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
library(sf)
```

Attempting to use High Park Fire in Larimer county as an example.

```{r}
# read county sf
larimer_boulder_county_sf <- st_read(dsn = "../../data/shapefiles/us_county/") %>% 
  filter(STATEFP == "08" & COUNTYFP %in% c("069", "013")) %>% 
  mutate(fips = paste0(STATEFP, COUNTYFP))

# extract crs
wgs84 <- st_crs(larimer_boulder_county_sf)

# read grid and filter to larimer grid
grid_sf <- st_read(dsn = "../../data/shapefiles/co_krig_grid/") %>% 
  filter(COLX %in% 123:129) %>% 
  filter(ROWY %in% 102:110)
# assign wgs84
st_crs(grid_sf) <- wgs84

ggplot() +
  geom_sf(data = grid_sf, fill = NA, color = "blue") +
  geom_sf(data = larimer_boulder_county_sf, fill = NA,  color = "red") +
  theme_minimal()
```

Grid intersect to find which grids belong to which counties.

```{r}
# find county intersects with grid id
grid_county_key <- st_intersection(larimer_boulder_county_sf, grid_sf) %>% 
  select(GRID_ID, fips, COLX, ROWY, NAME) %>% 
  mutate(GRID_ID = as.character(GRID_ID),
         county_name = as.character(NAME)) %>% 
  select(-NAME)
# remove geometry atribute
st_geometry(grid_county_key) <- NULL

larimer_key <- grid_county_key %>% 
  filter(ROWY > 104 & county_name == "Larimer")

boulder_key <- grid_county_key %>% 
  filter(ROWY <= 104 & county_name == "Boulder")

county_key <- bind_rows(larimer_key, boulder_key)
```


Loading grid values to get mean value and standard error for PM~2.5~ values in the grid during the summer.

```{r}
# read grid pm
grid_pm <- read_csv("../../data/smoke/1015-grid_pm.csv", 
                    col_types = "cDdddd") %>% 
  filter(GRID_ID %in% grid_county_key$GRID_ID) %>% 
  mutate(pm_diff_grid = pm25_grid - sbg_pm_grid,
         month = as.factor(lubridate::month(date)), # extract month as factor
         year = as.factor(lubridate::year(date)), # extract year as factor
         # month_day
         day_of_year = as.numeric(format(date, "%j")),
         season = as.factor(case_when(month %in% c(12, 1, 2) ~ "winter",
                                      month %in% c(3:5) ~ "spring",
                                      month %in% c(6:8) ~ "summer",
                                      month %in% c(9:11)~ "fall")),
         # set krig values lower than 0 to 0
         pm25_grid = if_else(pm25_grid < 0, 0, pm25_grid),
         smk5_g_hms = if_else(pm_diff_grid > 5 & 
                                month %in% c(4:10) & hms_grid == 1, 1, 0),
         # transforming pm kriged estimates to a 10 unit increase in pm
         pm25_g_10u = pm25_grid/10) %>% 
  # filter to summer of 2012 high park fire dates
  filter(season %in% c("summer", "fall")) %>% 
  left_join(county_key, by = "GRID_ID")
```

Plot of county time series of PM~2.5~ values.

```{r larimer_grid_pm_plot}
county_avg <- grid_pm %>% 
  group_by(county_name, day_of_year, year) %>% 
  summarise(pm25_mu = mean(pm25_grid), pm25_max = max(pm25_grid), 
            bg_mu = mean(sbg_pm_grid))


ggplot(county_avg, aes(x=day_of_year, y=pm25_max, 
                       group=county_name, color = county_name)) +
  geom_point(size = 0.2) +
  facet_wrap(~year) +
  theme_minimal()
```

High park fire started on June 9th, 2012 and was 100% contained on June 30, 2012. The PM~2.5~ peaks agree with this I think, showing the highest spike in air pollution probably around June 15th, and then again at the end of the month and in to early July.

HMS grid.

```{r}
hms_count <- grid_pm %>% 
  group_by(day_of_year, year) %>% 
  summarise(n = n(), hms_grid_count = sum(hms_grid)) %>% 
  mutate(smoke_prop = hms_grid_count/n)
  
ggplot(hms_count, aes(x=day_of_year, y=smoke_prop)) +
  geom_point() +
  facet_wrap(~year)
```

Interested in outcomes counts during this time for Larimer.

```{r cvd_resp_mortality}
# load casecross  mortality list
load("../../data/health/co_mortality_cc_list.RData") 
# load icd10 outcome list
load("../../data/health/icd10_outcome.RData")

# co cvd and resp mortaility
co_death_list <- casecross_list[c(1,4)]

# removing casecross_list to save memeory
rm(casecross_list)
```

Loading morbidity data.

```{r resp_cvd_morbidity}
# load casecross list
load("../../data/health/1015-co_morbidity_casecross_list.RData")
# load icd9
load("../..//data/health/icd9_outcome_vectors.RData")

# co hosp list
co_hosp_list <- co_morbidity_cc_list[c(1,6)]

# removing 
rm(co_morbidity_cc_list)
```

```{r mortality_rate}
# read in county population denominators for colorado
county_pop <- read_csv("../../data/health/2016-colorado_population.csv") %>% 
  rename(fips = GEO.id2) %>% 
  select(fips, respop72010:respop72015) %>% 
  filter(fips %in% c("08069","08013")) %>% 
  gather(year, population, -fips) %>% 
  mutate(year = as.numeric(str_sub(year, start = 8)))

# assign death outcome
death_outcome <- names(co_death_list)


# time series of weekly death and rate for cardiovascular and respiratory
death_ts <- map2(.x = co_death_list,  .y = death_outcome, 
                 ~ filter(.x, outcome == 1 & fips %in% c("08069","08013")) %>% 
                   mutate(date = as.Date(as.character(date_of_death)),
                          week = floor_date(date, unit = "week")) %>% 
                   group_by(week, fips) %>% 
                   summarise(n = n()) %>% 
                   left_join(county_names, by = "fips") %>% 
                   mutate(year = year(week),
                          month = lubridate::month(week), 
                          day_of_year = as.numeric(format(week, "%j")),
                season = as.factor(case_when(month %in% c(12, 1, 2) ~ "winter",
                                      month %in% c(3:5) ~ "spring",
                                      month %in% c(6:8) ~ "summer",
                                      month %in% c(9:11)~ "fall")),
                          outcome = .y) %>% 
                   left_join(county_pop, by = c("fips", "year")) %>% 
                   mutate(rate_per_100k = (n/population)*100000) %>% 
                   filter(week >= "2010-01-01" & week <= "2015-12-31")) %>% 
  map_dfr(., rbind)
```

Summer fall death plot.

```{r}
sum_fall_death_ts <- death_ts %>% 
  filter(season %in% c("summer", "fall")) %>% 
  filter(outcome == "cvd")

ggplot(sum_fall_death_ts, aes(x=day_of_year, y=rate_per_100k, 
                       group=county_name, color = county_name)) +
  geom_point(size = 0.2) +
  facet_wrap(~year) +
  theme_minimal()
```

Larimer county case-cross association.

```{r}
# load grid id key
grid_key <- read_csv("../../data/shapefiles/wrfgrid_key.csv", 
                     col_types = cols(.default = "d", GRID_ID = "c", 
                                      WRFGRID_ID = "c")) %>% 
  select(GRID_ID, WRFGRID_ID)


resp_deaths_larimer <- co_death_list$resp %>% 
  filter(fips == "08069") %>% 
  mutate(outcome = as.numeric(as.character(outcome)),
         date = as.Date(as.character(date_of_death)),
         WRFGRID_ID = as.character(wrfgrid_id)) %>% 
  filter(date > "2012-06-01" & date < "2012-09-01") %>% 
  left_join(grid_key, by = "WRFGRID_ID") %>% 
  left_join(grid_pm, by = c("date", "GRID_ID"))


mod <- clogit(outcome ~ pm25_g_10u*hms_grid + strata(id), data = resp_deaths_larimer)
summary(mod)
```



Now I'm interested in both Boulder County during the Four-Mile Canyon Fire and health outcomes.
