---
title: "Census tract cardiopulmonary rates"
author: "Ryan Gan"
date: "10/24/2017"
output: html_document
---

## Document background/overview

As part of the Colorado chapter of the American Lung Assocition's project to qunatify the health impacts of two coal-fired power plants in Colorado Springs and Pueblo, we need to calculate baseline cardiopulmonary emergency department (ED) visits in Colorado from 2010 to September of 2015. We've decided to limit the period of estimates from 2010 to 2014. I'm using the 2014 census estimates and shapefiles Sheena provided.

Set up. Calling tidyverse and sf packages.
```{r setup, echo=F, message=F, warning=F}
library(tidyverse)
library(sf)
```

Importing census denominators and simple features object as well as hospital claims data. 

```{r data import, echo=F, message=F, warning=F}
# read in census assignments
id_path <- "../data/smoke/colorado_census_ids.csv"
census_id <- read_csv(id_path) %>% 
  mutate(WRFGRID_ID = as.character(WRFGRID_ID))

# read in the dataframe with outcomes 
hosp_path <- "../data/health/co_hosp_w_outcome_df.csv"
co_hosp <- read_csv(hosp_path) %>% 
  # noticed broomfield county_geo 014 was sometimes assigned county_final 159
  # there is no county 159. All other spatial indicators suggest broomfield 014
  mutate(county_final = ifelse(county_geo == "014" & county_final == "159", 
                               "014", county_final),
    # make fips code
    FIPS = paste0("08", county_final),
    # set wrfgrid values of 0 to missing
    WRFGRID_ID = ifelse(WRFGRID_ID == 0, NA, WRFGRID_ID),
    WRFGRID_ID = as.character(WRFGRID_ID)) %>% 
  # remove those missing a WRFGRID ID
  filter(!is.na(WRFGRID_ID))

# join with poly ids
co_hosp_census <- co_hosp %>%
  # join with wrfgrid 
  left_join(census_id, by = "WRFGRID_ID") %>% 
  # mutate WRFGRID ID to factor
  mutate(WRFGRID_ID = as.factor(WRFGRID_ID),
  # adding in a season variable for estimation by season 
  season = as.factor(ifelse(ADMMM >= 3 & ADMMM <= 5, "spring",
             ifelse(ADMMM >= 6 & ADMMM <= 8, "summer",
               ifelse(ADMMM >= 9 & ADMMM <= 11, "fall", "winter")))),
  # adding in a year indicator for season to check for trends
  season_yr = as.factor(paste0(season, ADMYY)))

# import census denominators for counties and states
census_estimates <- read_csv("../data/health/co-est2016-alldata.csv") %>% 
  # make fips variable
  mutate(FIPS = paste0(STATE, COUNTY))

# lat/lon of power plants ----
plant_id <- c("pueblo", "colo_springs")
lat <- c(38.2081, 38.8244)
lon <- c(-104.5747, -104.8331)

plant_locations <- tibble(plant_id, lat, lon) 
```

Importing simple features that will be used for maps latter.
```{r simple features import, echo=F, warning=F, message=F}
# read shapefile Sheena sent 
co_tract_2014_path <- "../data/shapefiles/2014_CO_TRACT"
co_census_2014 <- st_read(dsn=co_tract_2014_path, layer="CO_Tracts_2014")
# save crs 
nad83 <- st_crs(co_census_2014)

# read county shapefile
county_path <- "../data/shapefiles/colorado_county"
county_sf <- st_read(dsn=county_path, layer = "colorado_county") %>% 
  # make fips variable
  mutate(FIPS = paste0(STATEFP, COUNTYFP))
# assign nad83
st_transform(county_sf, nad83)

# subset to study area 
# pikes peak study area: Teller and El Paso
# soutern central: Custer, Huerfano, Pueblo, and Las Animas
ala_study_sf <- county_sf %>% 
  filter(NAME %in% c("Teller", "El Paso", "Custer", "Huerfano", 
                   "Pueblo", "Las Animas")) %>% 
  st_transform(nad83)

# read in populations and metadata for the tracts
# i believe i may only need B00001e1 
census_data <- read_csv(paste0("../data/shapefiles/2014_CO_TRACT/",
  "Populations_CO_Tracts_2014.txt")) 

# limit census data to just total population
census_population <- census_data %>% 
  select(GEOID, B01001e1) %>% 
  # need to only extract the last part of the GEOID to join 
  mutate(GEOID = stringr::str_sub(GEOID, start = 8L)) %>% 
  rename(total_pop = B01001e1)
```
## Colorado trends

Evaluating state-wide trend in cardiopulmonary ED/urgent care hospital admissions. I will eventually limit to ED visits, but I want to show something weird. Creating a time-series for the entire state and calculating rates per 100,000 persons for cardiopulmonary, CVD, and respiratory ED or urgent care hospital visits. Population can vary by year, but I assume a yearly denominator is an appropriate denominator for days in the year. 

```{r state timeseries dataframe, echo=F, message=F, warning=F}
# subset out colorado census estimates
colorado_denom <- census_estimates %>% 
  filter(FIPS == "08000") %>% 
  # select variables with popestimate in title
  select_at(vars(contains("POPESTIMATE"))) %>% 
  gather(year, population) %>% 
  # split out only the year
  mutate(admit_year = as.numeric(stringr::str_replace(year,
    "POPESTIMATE", ""))) %>% 
  select(admit_year, population)

# time series
state_daily_count <- co_hosp_census %>% 
  # group by admit dat
  group_by(admit) %>% 
  summarise(cvd_n = sum(cvd_dx), resp_n = sum(resp_dx)) %>% 
  # make sum of cvd and resp outcomes and admit_year variable to join with denom
  mutate(cardiopulm_n = cvd_n + resp_n,
         admit_year = lubridate::year(admit)) %>% 
  left_join(colorado_denom, by = "admit_year") %>% 
  # calculate daily rates per 10,000 persons
  mutate(cardiopulm_per_100k = (cardiopulm_n/population)*100000,
         cvd_per_100k = (cvd_n/population)*100000,
         resp_per_100k = (resp_n/population)*100000) %>% 
  # gather rate columns in one column 
  gather("outcome", "rate", cardiopulm_per_100k:resp_per_100k)
```

Plotting daily rates over time. Trying a small-multiples so the patterns don't overlap. Looks like respiratory ED/urgent care admissions have the standard seasonal pattern going, which is driving the aggregated pattern in the cardiopulmonary ED/urgent care visits. CVD ED/urgent care vists don't really have a seasonal pattern. Across all three outcomes, there is a general decline in ED/urgent care admissions (added in a linear regression line just because). Interesting that in 2013 there is a dip around Christmas and the New Years, which I see sometimes, but also there is a big dip across all three outcomes near the end of the date range of my data, 2015-09-30; this may suggest that the dip is maybe more of a QC issue where admissions haven't been reconciled by this time?

```{r colorado daily rate plots, echo=F, message=F, warning=F}
# plot of rate
p_rate <- ggplot(data = state_daily_count, 
                 aes(x=admit, y=rate, group=outcome, colour=outcome)) +
  geom_point(size = 0.25) +
  facet_wrap(~outcome) +
  stat_smooth(method = "glm", colour = "black", fill = "black", linetype=2) +
  ylab("Daily rate per 100k persons") +
  xlab("Date") +
  ggtitle("Rate of cardiopulmonary ED/urgent care visits in Colorado") +
  theme_bw()
# plot chart
p_rate
```

### Limiting to ED visits

Subsetting to only ED visits as most concentration-response functions are only for ED visits. 

```{r ed timeseries, echo=F, message=F, warning=F}
ed_hosp <- co_hosp_census %>% 
  # filter to only ED visits
  filter(ADMSNO == 1)

# time series
state_daily_count <- ed_hosp %>% 
  # group by admit dat
  group_by(admit) %>% 
  summarise(cvd_n = sum(cvd_dx), resp_n = sum(resp_dx)) %>% 
  # make sum of cvd and resp outcomes and admit_year variable to join with denom
  mutate(cardiopulm_n = cvd_n + resp_n,
         admit_year = lubridate::year(admit)) %>% 
  left_join(colorado_denom, by = "admit_year") %>% 
  # calculate daily rates per 10,000 persons
  mutate(cardiopulm_per_100k = (cardiopulm_n/population)*100000,
         cvd_per_100k = (cvd_n/population)*100000,
         resp_per_100k = (resp_n/population)*100000) %>% 
  # gather rate columns in one column 
  gather("outcome", "rate", cardiopulm_per_100k:resp_per_100k)
```

ED rate only plot.

```{r colorado ed daily rate plots, echo=F, message=F, warning=F}
# plot of rate
p_rate <- ggplot(data = state_daily_count, 
                 aes(x=admit, y=rate, group=outcome, colour=outcome)) +
  geom_point(size = 0.25) +
  facet_wrap(~outcome) +
  stat_smooth(method = "glm", colour = "black", fill = "black", linetype=2) +
  ylab("Daily rate per 100k persons") +
  xlab("Date") +
  ggtitle("Rate of cardiopulmonary EDvisits in Colorado") +
  theme_bw()
# plot chart
p_rate
```

There is a weird cluster of low ED visit rates in the first part of 2010. When urgent care was included, I didn't notice it.It suggests to me that something happend systematically. I could handle this a couple different ways. I could include both ED and urgent care visits, but the rate may be a little higher on average than ED only. This may not be appropriate for an HIA where most other estimates will be ED visit only.

Two other more desirable options would be to exclude this winter/spring period in 2010, or use emprical Bayesian weighting to estimate the posterior using the average/distribution rate of the entire period as the prior.


### Seasonal rates

State-wide seasonal rate over the time period. The idea behind this would be to account for different seasonal rates (particularly spring and fall) when pollutant concentrations from the plants may be different. I believe the exposure folks were thinking of estimating a representative spring and fall month.

```{r state seasonal trends, echo=F, message=F, warning=F}
seasonal_rates <- ed_hosp %>% 
  group_by(ADMYY, season) %>% 
  summarise(cvd_n = sum(cvd_dx), resp_n = sum(resp_dx)) %>% 
  # make sum of cvd and resp outcomes and admit_year variable to join with denom
  mutate(cardiopulm_n = cvd_n + resp_n,
         admit_year = as.numeric(paste0(20,ADMYY))) %>% 
  left_join(colorado_denom, by = "admit_year") %>% 
  # calculate daily rates per 100k persons
  mutate(cardiopulm_per_1000py = (cardiopulm_n/population)*1000,
         cvd_per_1000py = (cvd_n/population)*1000,
         resp_per_1000py = (resp_n/population)*1000) %>% 
  # gather rate columns in one column 
  gather("outcome", "rate", cardiopulm_per_1000py:resp_per_1000py) %>% 
  # remove 2015 as it's not a full year of data
  filter(admit_year != 2015)
```

Plot of seasonal rate.I calculated rate per population for each 3-month season across the state. We could think about averaging the rates. I tried a rate over the season for the 5 year period, but I need to think about how this is appropriate for

```{r plot seasonal rate state, echo=F, message=F, warning=F}
# plot of rate
p_rate <- ggplot(data = seasonal_rates, 
    aes(x=season, y=rate, group=as.factor(admit_year), 
        colour=as.factor(admit_year))) +
  geom_point(size = 0.5) +
  facet_wrap(~outcome) +
  ylab("Seasonal rate per 1000 persons per 3 months") +
  xlab("Season") +
  ggtitle("Rate of cardiopulmonary ER/urgent care visits in Colorado") +
  theme_bw()
# plot chart
p_rate
```

Spring and winter of 2010 are definately outliers. CVD rates don't really look that different by years for seasons. I don't really like this plot. I want to work in a time scale (YYYY-mid month of season) but I'll come back to it. 

### Aggregated seasonal rates

Calculating the rate for each season over a five year period. Still trying to grappel with this concept. Bring it up with Sheryl and Sheena for discussion. 

### Colorado period estimate 2010 to 2014

I have a count of outcomes over the 2010-2014 5 year period, and a denominator of persons in various spatial units over this time. I want to estimate the rate of outcomes per X persons per 5 years. I think I need to assume that each person will contribute a 5 years of time, so I need to multiple they denominator by 5. Discuss this assumption with Sheena and Sheryl. 

Estimating the rate for the entire state over the study period.

```{r state period estimates, echo=F, message=F, warning=F}
state_rate_period <- ed_hosp %>% 
  # filter to period
  filter(ADMYY != 15) %>% 
  # summarise
  summarise(cvd_n = sum(cvd_dx), resp_n = sum(resp_dx)) %>% 
  mutate(cardiopulm_n = cvd_n + resp_n,
    # cacluating population using old standard of beginning + end population /2
    population = (5048644+5349648)/2,
    person_time = population*5,
    cardiopulm_per_1000p5y = (cardiopulm_n/person_time)*1000,
    cvd_per_1000p5y = (cvd_n/person_time)*1000,
    resp_per_1000p5y = (resp_n/person_time)*1000)

# rate 
state_rate_period
```

I think the rate should be interpreted as the number of events per 100 persons over the 5 year period. I assume that each person in the denominator is contributing five years of person time, so I multiple the denominator by 5. This assumption is open for debate.

Trying to calculate spring, summer, fall, winter seasonal rates of 15 month chunks.

```{r seasonal 15 month rate, echo=F, message=F, warning=F}
# sum up seasonal estimates and calculate rate
season_agg_rates <- ed_hosp %>% 
  filter(ADMYY != 15) %>% 
  group_by(season) %>% 
  summarise(cvd_n = sum(cvd_dx), resp_n = sum(resp_dx)) %>% 
  # make sum of cvd and resp outcomes and admit_year variable to join with denom
  mutate(cardiopulm_n = cvd_n + resp_n,
         # 2010 + 2014 pop divided by 2
         population = 5199146,
         # 15 months is my time period so multiplying denom by 1.25
         person_time = population * 1.25,
         cardiopulm_per_1000py = (cardiopulm_n/person_time)*1000,
         cvd_per_1000py = (cvd_n/person_time)*1000,
         resp_per_1000py = (resp_n/person_time)*1000)
```

## County trends

Examining period cardiopulmonary rates by Colorado county. I didn't do a time-series because there were a lot of 0 values at dates, leading to many of the less populated counties with unstable rates at the daily level.

```{r county rate dataframe, echo=F, message=F, warning=F}
# denom of 2010+2014/2 for each county
county_denom <- census_estimates %>% 
  filter(STATE == "08" & COUNTY != "000") %>% 
  # select variables with popestimate in title
  select(FIPS, POPESTIMATE2010, POPESTIMATE2014) %>% 
  # split out only the year
  mutate(population = round((POPESTIMATE2010+POPESTIMATE2014)/2,0)) %>% 
  select(FIPS, population)

# read colo geofacet grid
colo_grid <- read_csv("../data/shapefiles/colo_county_grid.csv")

# county name of fips
county_name <- colo_grid %>% select(code, name) %>% 
  rename(county_name = name, FIPS = code)

# time series
county_rate <- ed_hosp %>% 
  # filter to period
  filter(ADMYY != 15) %>% 
  # group by county fips
  group_by(FIPS) %>% 
  summarise(cvd_n = sum(cvd_dx), resp_n = sum(resp_dx)) %>% 
  # make sum of cvd and resp outcomes and admit_year variable to join with denom
  mutate(cardiopulm_n = cvd_n + resp_n) %>% 
  left_join(county_denom, by = c("FIPS")) %>% 
  # calculate daily rates per 10,000 persons
  mutate(person_time = population*5,
    cardiopulm_per_1000p5y = (cardiopulm_n/person_time)*1000,
    cvd_per_1000p5y = (cvd_n/person_time)*1000,
    resp_per_1000p5y = (resp_n/person_time)*1000)
```

Maps of the period rate of cardiovascular and pulmonary ED visits by county.

County-level cardiovascular rate per 1000 persons from 2010 to 2014.

```{r cvd period county rate maps, echo=F, message=F, warning=F}
# join county rates to county sf object
county_rate_sf <- county_sf %>% 
  left_join(county_rate, by = "FIPS")

# cvd plot
cvd_p <- ggplot(county_rate_sf) +
  geom_sf(aes(fill=cvd_per_1000p5y), colour = "white") +
  scale_fill_gradient(name = "CVD Rate",low = "#cbb4d4", high = "#20002c") +
  ggtitle("County CVD ED rates per 1000 persons from 2010 to 2014") +
  theme_bw()
# print map
cvd_p
```

County-level respiratory rate per 1000 persons from 2010 to 2014.
```{r resp period county rate maps, echo=F, message=F, warning=F}
# resp plot
resp_p <- ggplot(county_rate_sf) +
  geom_sf(aes(fill=resp_per_1000p5y), colour = "white") +
  scale_fill_gradient(name = "Resp Rate",low = "#26d0ce", high = "#1A2980") +
  ggtitle("County respiratory ED rates per 1000 persons from 2010 to 2014") +
  theme_bw()
# print map
resp_p
```

Rates look higher in Pueblo county.

## Census Estimates

Estimating period rates for census tracts in Colorado and then subsetting to the ALA study area to make viewing easier.

```{r census period rate, echo=F, message=F, warning=F}
census_rate_period <- co_hosp_census %>% 
  filter(ADMYY != 15) %>% 
  group_by(GEOID) %>% 
  summarise(cvd_n = sum(cvd_dx), resp_n = sum(resp_dx)) %>% 
  # join with census denom
  left_join(census_population, by = "GEOID") %>% 
  # calculate rate 
  mutate(cardiopulm_n = cvd_n + resp_n,
  total_pop5y = total_pop*5,
  cardiopulm_per_100p5y = (cardiopulm_n/total_pop5y)*100,
  cvd_per_100p5y = (cvd_n/total_pop5y)*100,
  resp_per_100p5y = (resp_n/total_pop5y)*100)

# head of these calculations
head(census_rate_period)

# join with colorado census 2014
census_rate_sf <- co_census_2014 %>% 
  mutate(GEOID = as.character(GEOID)) %>% 
  left_join(census_rate_period, by = "GEOID")
```

Although I calculated rates for the entrie state, I'm going to subset to the study area to make it more visible.

```{r ala cvd rate map, echo=F, message=F, warning=F}
# ALA study area
ala_sf <- st_intersection(census_rate_sf, ala_study_sf)

# plot
p <- ggplot(ala_sf) +
  geom_sf(aes(fill=cvd_per_100p5y), colour = "white") +
  scale_fill_gradient(name="CVD Rate", low = "#cbb4d4", high = "#20002c") +
  geom_sf(data=ala_study_sf, alpha = 0) +
  geom_point(data=plant_locations, aes(x=lon,y=lat), colour= "red") +
  ggtitle("ALA census CVD ED rates per 100 persons from 2010 to 2014") +
  theme_bw()
# print map
p
```

Census respiratory rate map.
```{r ala resp rate map, echo=F, message=F, warning=F}
# plot
p <- ggplot(ala_sf) +
  geom_sf(aes(fill=resp_per_100p5y), colour = "white") +
  scale_fill_gradient(name= "Resp Rate", low = "#26d0ce", high = "#1A2980") +
  geom_sf(data=ala_study_sf, alpha = 0) +
  geom_point(data=plant_locations, aes(x=lon,y=lat), colour= "red") +
  ggtitle("ALA census respiratory ED rates per 100 persons from 2010 to 2014") +
  theme_bw()
# print map
p
```

Limiting to El Paso to view Colorado Springs area.

```{r el paso cvd map, echo=F, message=F, warning=F}
el_paso_sf <- ala_study_sf %>% 
  filter(NAME == "El Paso") %>% 
  st_transform(nad83)

el_paso_census_sf <- st_intersection(census_rate_sf, el_paso_sf)

# cs plant
cs_plant <- plant_locations %>% filter(plant_id == "colo_springs")

# plot
p <- ggplot(el_paso_census_sf) +
  geom_sf(aes(fill=cvd_per_100p5y), colour = "white") +
  scale_fill_gradient(name = "CVD Rate", low = "#cbb4d4", high = "#20002c") +
  geom_sf(data=el_paso_sf, alpha = 0) +
  geom_point(data=cs_plant, aes(x=lon,y=lat), colour= "red") +
  ggtitle("El Paso CVD ED rates per 100 persons from 2010 to 2014") +
  theme_bw()
# print map
p
```

Respiratory ED rate.

```{r resp cvd map, echo=F, message=F, warning=F}
# plot
p <- ggplot(el_paso_census_sf) +
  geom_sf(aes(fill=resp_per_100p5y), colour = "white") +
  scale_fill_gradient(name = "Resp Rate", low = "#26d0ce", high = "#1A2980") +
  geom_sf(data=el_paso_sf, alpha = 0) +
  geom_point(data=cs_plant, aes(x=lon,y=lat), colour= "red") +
  ggtitle("El Paso CVD ED rates per 100 persons from 2010 to 2014") +
  theme_bw()
# print map
p
```

Missing a lot of estimates for Colorado Springs census tracts. 

### Thoughts on unstable rates

I considered small area estimation, but I would need the census survey dataframe and a variable of small area (census tract) in the dataframe to estimate expected population? in that tract. I don't think it would work in this case because I'm trying to use aggregated estimates of a larger area to estimate rates in a smaller area. I still think emprical bayes might be a good approach here.

