---
title: "Descriptive Epi for Sheryl's Grant"
author: "Ryan Gan"
date: '2018-02-09'
output:
  html_document:
    df_print: paged
editor_options:
  chunk_output_type: console
---

## Introduction

Descriptive cardiopulmonary hospital inpatient numbers for Colorado for Sheryl's grant application. Note that these data only contain billing records for inpatient hosptial stays from 2010 to September 30th, 2015. They only contain billing records for cardiopulmonary outcomes.

Setting up chunk options for markdown output. Hidding code messages, warning messages, and code chunks. Calling packages for this document.

```{r setup}
# knitr options
knitr::opts_chunk$set(fig.height = 6, fig.width = 8, fig.align = "center",
                      progress = F, message = F, warning = F, echo=F,
                      results = "asis")

# libraries
library(tidyverse)
library(sf)
library(raster)
library(ncdf4)
```

Importing Colorado cardiopulmonary inpatient data, coding for respiratory outcomes, and filtering to only primary diagnosis of respiratory outcomes. Note, I'm able to seperate out unique visits, but not unique persons with the provided data. Also, I do not have outcomes after 2015-09-30 because of the transition from ICD-9 to ICD-10 nationwide.

```{r hospital_data}
# load outcomes list
load("../data/health/icd9_outcome_vectors.RData")

# define relative path to data folder
co_path <- paste0("../data/health/co_hosp_1015.csv")
# read in dataframe using read_csv function from tidyverse package
co_resp_hosp <- read_csv(co_path, progress = F, 
                         col_types = cols(.default = "c")) %>% 
 filter(ADMTNO == 1) %>% # subset to ADMTNO == 1 Emergency
  mutate(date = as.Date(admit, format = "%m/%d/%Y"),
    resp = if_else(dx1 %in% icd9_outcomes$resp, 1, 0),
    asthma = if_else(dx1 %in% icd9_outcomes$asthma, 1, 0),
    copd = if_else(dx1 %in% icd9_outcomes$copd, 1, 0),
    sex_factor = case_when(SEX == "1" ~ "M",
                           SEX == "2" ~ "F"),
    # age as numeric; if any age is greater than 120, set to missing
    age = ifelse(as.numeric(AGEYRS) > 120, NA, as.numeric(AGEYRS)),
    age_cat = as.factor(case_when(age <= 4 ~ "0-4",
                        age > 4 & age <= 14 ~ "4-14",
                        age > 14 & age <= 44 ~ "15-44",
                        age > 44 & age <= 64 ~ "44-64",
                        age > 64 & age <= 84 ~ "65-84",
                        age > 85 ~ "85+")),
    fips = paste0("08", county_final),
    race_cat = as.factor(case_when(RACE == "1" ~ "white",
                         RACE == "2" ~ "black",
                         RACE == "3" ~ "asian",
                         RACE == "4" ~ "native_american",
                         RACE == "5" ~ "hispanic",
                         RACE == "6" ~ "other")),
    year = lubridate::year(date)) %>% 
  # filter to respiratory outcomes
  filter(resp == 1)
```

## Colorado Estimates

I want to understand some aggregate trends for the state just so I know if my study area-specific numbers are reasonable. Summing up descriptive epidemiologic estimates for the State of Colorado from 2010-01-01 to 2015-09-30.

### Observations by Year

Counting up total observations for each year. Note, 2015 will be smaller since it's only 9 months.

```{r year_counts}
n_obs <- co_resp_hosp %>% 
  group_by(year) %>% 
  summarise(n_obs = n())

n_obs
```


### Age Distributions by Year

Most inpatient observations admitted through the emergency room are in young childre under the age of 10 and in those 65 and older. I don't really see a difference in distribution of age by year. I have 16 observations that were either missing age (n=1), or not real (n=15 in the 800s). I could not tell if perhaps an extra 0 was added, and in the end decided to set these observations to missing.

```{r age_dist_plot}
ggplot(data=co_resp_hosp, aes(x=age)) +
  geom_density() +
  facet_wrap(~year)
```

### Sex Distribution
Counts and proportion by year and sex. F=Female, M=Male, _prop indicates row proportion. Respriatory inpatient hospitalizations via ER is proportionally similar between males and females over the 6 year period.
```{r sex_distribution}
sex_tab <- co_resp_hosp %>% 
  group_by(year, sex_factor) %>%
  summarise(n_obs = n()) %>% 
  spread(sex_factor, n_obs) %>% 
  mutate(n_obs = (F+M),
    F_prop = round(F/n_obs,2),
    M_prop = round(M/n_obs,2)) %>% 
  dplyr::select(year, F, F_prop, M, M_prop)

sex_tab
```

### Race/Ethnicity Distribution

Presenting total respiratory inpatient admitted through ER observations for each year and then proportion coded for each race by year. Mostly white (~70%), followed by hispanic and black. Quite a bit of missing data too, which is common with claims data.

```{r race_tab}
race_tab <- co_resp_hosp %>% 
  group_by(year, race_cat) %>% 
  summarise(n = n()) %>% 
  mutate(race_cat = as.factor(ifelse(is.na(race_cat), 
                      "missing", as.character(race_cat)))) %>% 
  spread(key = race_cat, value = n) %>% 
  left_join(n_obs, by = "year") %>% 
  # proportion
  mutate_at(vars(asian:white), funs(prop = round(./n_obs,2))) %>% 
  dplyr::select(year, n_obs, white, white_prop, black, black_prop, hispanic,
         hispanic_prop, asian, asian_prop, native_american,
         native_american_prop, other, other_prop, missing, missing_prop)

race_tab
```
 
### Respiratory Outcome Distributions
Breaking down yearly counts of respiratory observations by asthma and COPD.
```{r respiratory_outcomes}
outcome_tab <- co_resp_hosp %>% 
  mutate(outcome = as.factor(case_when(asthma == 1 ~ "asthma",
    copd == 1 ~ "copd",
    resp == 1 & asthma == 0 & copd == 0 ~ "resp_other"))) %>% 
  group_by(year, outcome) %>% 
  summarise(n = n()) %>% 
  spread(key="outcome", value = "n") %>% 
  mutate(n_obs = (asthma + copd + resp_other)) %>% 
  mutate_at(vars(asthma:resp_other), funs(prop=round(./n_obs,2))) %>% 
  dplyr::select(year, n_obs, asthma, asthma_prop, copd, copd_prop, 
         resp_other, resp_other_prop)

outcome_tab
```

Now that I have some state-wide estimates, I'l evaluate numbers for the front range.

## Grant Numbers: Colorado Front Range Study Area

The following estimates are likely going to be of more use for the grant.

Starting with a study area map.

```{r bbox_clip_function}
# clip by bbox function ------
bbox_clip <- function(sf, bbox) {
  # find the CRS of the sf object
  crs <- sf::st_crs(sf)$proj4string
  # create matrix
  x <- c(bbox[1], bbox[1], bbox[3], bbox[3], bbox[1])
  y <- c(bbox[2], bbox[4], bbox[4], bbox[2], bbox[2])
  coords <- matrix(cbind(x, y), ncol=2)
  # create polygon and assign same coord crs as sf object
  coords_poly <- sp::Polygon(coords)
  bbox_poly <- sp:: SpatialPolygons(list(sp::Polygons(list(coords_poly),
    ID = "bbox")), proj4string = sp::CRS(crs))
  # convert to sf feature
  bbox_sf <- st_as_sf(bbox_poly)
  # clip sf object
  clipped_sf <- sf[bbox_sf,]
  return(clipped_sf)
}

# clipping bounding box
study_bbox <- st_bbox(c(xmin=-105.3, xmax=-104.5, ymax=41, ymin = 38))
```

I believe Sheryl is intersted in the front-range population. I'll define it as along the I-25 corridor from the Wyoming border (maybe Ft. Collins) to Pueblo. For simplicity, I've subsetted the WRFGRID box to latitude 38 to 41 degrees N, and longitude 104.5 to 105.3 degrees W.

```{r county_sf}
# county subset
county_sub_name <- c("Larimer", "Weld", "Boulder", "Broomfield", "Adams", 
                "Denver", "Jefferson", "Arapahoe", "Douglas", "El Paso",
                "Pueblo")

# read in county shapefile and subset to only colorado fips
co_county_sf <- st_read("../../meta_wildfire/data/shapefile/us_county", 
                        layer = "us_county") %>%
  # limit to colorado
  filter(STATEFP == "08") %>% 
  # create fips variable
  mutate(fips = paste0(STATEFP, COUNTYFP))

# save wgs84 crs
wgs <- st_crs(co_county_sf)

# county clip
county_clip <- bbox_clip(co_county_sf, study_bbox)

# extract county names
county_text <- county_clip %>% 
  st_transform(wgs) %>% 
  st_centroid() %>% 
  st_coordinates() %>% 
  cbind(., as.character(county_clip$NAME)) %>%
  as_data_frame() %>% 
  rename(lon = X, lat = Y, county = V1) %>% 
  mutate(lon = as.numeric(lon), lat = as.numeric(lat))
```

Reading in I-25 and I-70 simple features for reference.

```{r interstate_sf}
# read in colorado roads shapefile
interstate <- st_read(paste0("../../meta_wildfire/data/",
  "shapefile/tl_2015_08_prisecroads"), layer = "tl_2015_08_prisecroads") %>% 
  # filter to I25 lines
  filter(RTTYP == "I") %>% 
  st_transform(crs = wgs)

# interstate clip
i_clip <- bbox_clip(interstate, study_bbox)
```

Reading in city boundary simple features.
```{r city_sf}
# read in city polygons 2010
city <- st_read("../data/shapefiles/Colorado_City_Point_Locations/", 
                layer = "Colorado_City_Point_Locations") %>% 
  st_transform(crs = wgs)

# limit cities 
city_points <- city %>% 
  filter(NAME %in% c("FORT COLLINS", "PUEBLO", "GREELEY", "BOULDER", "DENVER",
                     "COLORADO SPRINGS")) %>% 
  mutate(city = stringr::str_to_title(NAME))
```

```{r grid_sf}
# read wrfgrid shape
co_wrf_grid <- st_read(dsn = "../data/shapefiles/co_grid_shp", 
                       layer = "co_grid") %>% 
  st_set_crs(wgs) 

# clip grid box to study area
study_grid <- bbox_clip(co_wrf_grid, study_bbox)
```

Reading 2015 population estimate geotiff for Colorado. I made this file for the ALA project from the SEDAC 2015 global estimate.

```{r population_raster}
# read colorado 2015 population raster
co_pop_2015 <- raster("../data/shapefiles/2015-ColoradoPopDensity.tif")
# extract bbox of clipped county subset
fr_extent <- extent(st_bbox(county_clip)[c(1,3,2,4)])
# raster
fr_pop_2015 <- crop(co_pop_2015, fr_extent)

# i'm going to create a shapefile/simple features; easier to plot
front_range_pop_sf <- st_as_sf(rasterToPolygons(fr_pop_2015)) %>% 
  rename(popden = X2015.ColoradoPopDensity) %>% 
  # filtering to cells > 100 
  filter(popden > 100)

# cut once more to county shapefile so that populations outside the counties are not present
pop_clip <- front_range_pop_sf[county_clip,]
```

### Map of Study Area

Counties in the Colorado front-range area. Struggling with the legend.

Note: Sheryl wants population density instead of city boundaries. Also wants a black and white color scheme; or at least a map that can be printed easier. Maybe a reference map. I limited the population density to cells over 100 persons per square mile so that the sf file was smaller and easier to plot. It still takes a bit of time though.

Question for Sheryl: Do you actually want to see the grid or can it just be a bounding box?

Inset map.
```{r state_map}
# extent of county clip
study_bbox <- st_bbox(county_clip)[c(1,3,2,4)]
study_extent <- as_data_frame(matrix(study_bbox,
                       nrow = 1, ncol = 4, byrow = T)) %>% 
  rename(xmin = V1, xmax = V2, ymin = V3, ymax = V4)
# create county inset map
inset_map <- ggplot(co_county_sf) +
  geom_sf(fill = "transparent", color = "black", size = 0.1) +
  geom_rect(data = study_extent, 
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax), 
    fill = "red", color = "transparent", alpha = 0.5) +
  ggtitle("Colorado: Study Area") +
  theme(plot.title = element_text(size = 12),
        panel.background = element_blank(),
        panel.grid.major = element_line(colour = 'transparent'),
        panel.grid.minor = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        plot.background = element_rect(fill = "transparent", 
                                       color = "transparent"))
inset_map
```

Study map.
```{r study_map}
# plot map
study_map <-ggplot(pop_clip) +
  # start with plot of simple features of populations
  geom_sf(aes(fill=popden), color = NA) +
  # define aesthetics of poipulation
    scale_fill_gradient(name = expression("Population per km"^2), 
      low = "#26d0ce", high = "#1a2980") +
  # plot lines of counties
  geom_sf(data = county_clip, color = "black", 
          fill = "transparent", size = 0.5) +
  # plot i-25 and i-70
  geom_sf(data = i_clip, aes(color = "Interstate"), show.legend = "line") +
  # plot study grid
  geom_sf(data = study_grid, aes(color = "Study Grid"), fill = "transparent", 
          size = 0.1, show.legend = "line") +
  # custom colors for interstate and study grid
  scale_color_manual(values = c("Interstate" = "#0f9b0f", 
                                "Study Grid" = "red"), 
    labels = c("Interstate", "Study Grid"),
    name = "Boundary") + 
  # major city points and names
  geom_point(data = city_points, aes(x = LONG, y = LAT), color = "#3f2b96") +
  geom_text(data = city_points, aes(x = LONG, y = LAT, label = city), 
             color = "#3f2b96", size = 4, hjust = 1, vjust = -0.6) +
  # theme
  theme(panel.background = element_blank(),
        panel.grid.major = element_line(colour = 'transparent'),
        panel.grid.minor = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        axis.title = element_blank(),
        legend.key = element_rect(fill = NA, colour = NA, size = 0.25))

study_map
```

Final map with inset. Using the grid package.I will need to see if I can't figure this out a little better.
```{r map_w_inset}
# save map
# tiff(filename = "study_map.tiff", w=850, h)
grid::grid.newpage()
vpmain <- grid::viewport(width = 1, height = 1, x = 0.5, y = 0.5)
vpinset <- grid::viewport(width = 0.25, height = 0.25, x = 0.7, y = 0.85)
print(study_map, vp = vpmain)
print(inset_map, vp = vpinset)
```

Subsetting Colorado respiratory inpatient via ER observations to observations in the front range WRF-grid box. Note that approximately 79% of inpatient ER observations were in the front range area defined here. 

```{r resp_subset}
# grid id
wrf_id <- study_grid %>% 
  dplyr::select(WRFGRID_ID) 
# null geom
st_geometry(wrf_id) <- NULL

co_resp_fr <- co_resp_hosp %>% 
  filter(WRFGRID_ID %in% wrf_id$WRFGRID_ID) %>% 
  mutate(outcome = as.factor(case_when(asthma == 1 ~ "asthma",
    copd == 1 ~ "copd",
    resp == 1 & asthma == 0 & copd == 0 ~ "resp_other"))) 
```

### Total Number of Respiratory Inpatient Admitted via ER
Number of the whole Colorado dataset. Study period is 2010 to end of September 2015.

#### Overall n for Colorado
```{r co_n}
n_co <- base::nrow(co_resp_hosp)
n_co
```

#### Front Range n for Colorado
```{r fr_n}
n_obs <- base::nrow(co_resp_fr)
n_obs
```

#### Proportion of Front Range/Colorado
```{r fr_prop}
prop <- round(base::nrow(co_resp_fr)/base::nrow(co_resp_hosp),2)
prop
```

Estimating respiratory outcomes by year, sex, and age categories. I can use this dataset to estimate the numbers I need.

```{r front_range_strata_counts}
county_resp_obs <- co_resp_fr %>% 
  group_by(year, outcome, sex_factor, age_cat) %>% 
  summarise(n = n())
```

### Respiratory Outcome Observations by Year

Aggregating across all sub-groups to get counts of the outcomes by year for the front range subset. Proportions of total n observations are indicated by _prop.

```{r outcome_by_year}
outcome_counts_fr <- county_resp_obs %>% 
  group_by(year, outcome) %>% 
  summarise(n_obs = sum(n)) %>% 
  spread(outcome, n_obs) %>% 
  mutate(n_obs = (asthma+copd+resp_other)) %>% 
  mutate_at(vars(asthma:resp_other), funs(prop=round(./n_obs,2))) %>% 
  dplyr::select(year, n_obs, asthma, asthma_prop, copd, copd_prop, 
         resp_other, resp_other_prop)

outcome_counts_fr
```

### Respiratory Outcome Observations by Age Categories 

Aggregating outcome counts by age category. I would not really believe that the 0-4 or 4-14 age groups have any cases of COPD. These counts in this age group are probably because I have to use ICD-9 codes for other codes.

```{r fr_age}
age_counts <- county_resp_obs %>% 
  mutate(age_cat = forcats::fct_relevel(age_cat, 
    c("0-4", "4-14", "15-44", "44-64", "65-84", "85+"))) %>% 
  group_by(age_cat, outcome) %>% 
  summarise(n_obs = sum(n)) %>% 
  filter(!is.na(age_cat)) %>% 
  spread(outcome, n_obs) %>% 
  mutate(n_obs = (asthma+copd+resp_other)) %>% 
  mutate_at(vars(asthma:resp_other), funs(prop=round(./n_obs,2))) %>% 
    dplyr::select(age_cat, n_obs, asthma, asthma_prop, copd, copd_prop, 
          resp_other, resp_other_prop)

age_counts
```

### Respiratory Outcome Observations by Sex 
```{r fr_sex}
sex_counts <- county_resp_obs %>% 
  group_by(sex_factor, outcome) %>% 
  summarise(n_obs = sum(n)) %>% 
  filter(!is.na(sex_factor)) %>% 
  spread(outcome, n_obs) %>% 
  mutate(n_obs = (asthma+copd+resp_other)) %>% 
  mutate_at(vars(asthma:resp_other), funs(prop=round(./n_obs,2))) %>% 
    dplyr::select(sex_factor, n_obs, asthma, asthma_prop, copd, copd_prop, 
          resp_other, resp_other_prop)

sex_counts
```

### Respiratory Outcome by Sex and Age
```{r fr_age_sex}
age_sex_counts <- county_resp_obs %>% 
  mutate(age_cat = forcats::fct_relevel(age_cat, 
    c("0-4", "4-14", "15-44", "44-64", "65-84", "85+"))) %>% 
  group_by(age_cat, sex_factor, outcome) %>% 
  summarise(n_obs = sum(n)) %>% 
  filter(!is.na(age_cat)) %>% 
  spread(outcome, n_obs) %>% 
  mutate(n_obs = (asthma+copd+resp_other)) %>% 
  mutate_at(vars(asthma:resp_other), funs(prop=round(./n_obs,2))) %>% 
    dplyr::select(age_cat, sex_factor, n_obs, asthma, asthma_prop, copd, copd_prop, 
          resp_other, resp_other_prop)

age_sex_counts
```

Sheryl may want marginal proportions of age/sex strata that would be more informative for epi purposes. I'll present each outcomes where sex is the column header and age categories are the row headers.

### Asthma by Age and Sex
```{r asthma_age_sex}
asthma_as_tab <- county_resp_obs %>% 
  mutate(age_cat = forcats::fct_relevel(age_cat, 
    c("0-4", "4-14", "15-44", "44-64", "65-84", "85+"))) %>% 
  filter(outcome == "asthma") %>% 
  group_by(age_cat, sex_factor) %>% 
  summarise(n_obs = sum(n)) %>% 
  filter(!is.na(age_cat)) %>% 
  spread(sex_factor, n_obs) %>% 
  mutate(n_obs = (M+F)) %>% 
  mutate_at(vars(M:F), funs(prop=round(./n_obs,2))) %>% 
    dplyr::select(age_cat, n_obs, M, M_prop, F, F_prop)

asthma_as_tab
```
### COPD by Age and Sex
```{r copd_age_sex}
copd_as_tab <- county_resp_obs %>% 
  mutate(age_cat = forcats::fct_relevel(age_cat, 
    c("0-4", "4-14", "15-44", "44-64", "65-84", "85+"))) %>% 
  filter(outcome == "copd") %>% 
  group_by(age_cat, sex_factor) %>% 
  summarise(n_obs = sum(n)) %>% 
  filter(!is.na(age_cat)) %>% 
  spread(sex_factor, n_obs) %>% 
  mutate(n_obs = (M+F)) %>% 
  mutate_at(vars(M:F), funs(prop=round(./n_obs,2))) %>% 
    dplyr::select(age_cat, n_obs, M, M_prop, F, F_prop)

copd_as_tab
```
### Other Respiratory by Age and Sex
```{r other_resp_age_sex}
oresp_as_tab <- county_resp_obs %>% 
  mutate(age_cat = forcats::fct_relevel(age_cat, 
    c("0-4", "4-14", "15-44", "44-64", "65-84", "85+"))) %>% 
  filter(outcome == "resp_other") %>% 
  group_by(age_cat, sex_factor) %>% 
  summarise(n_obs = sum(n)) %>% 
  filter(!is.na(age_cat)) %>% 
  spread(sex_factor, n_obs) %>% 
  mutate(n_obs = (M+F)) %>% 
  mutate_at(vars(M:F), funs(prop=round(./n_obs,2))) %>% 
    dplyr::select(age_cat, n_obs, M, M_prop, F, F_prop)

oresp_as_tab
```

### All Respiratory Observations by Race
Numbers and proportions of racial categories. Total column is the total number of inpatient hospitalizations admitted via ER observations in the front range. This is for NIH enrollment numbers. 
```{r fr_race}
race_tab <- co_resp_fr %>% 
  group_by(race_cat) %>% 
  summarise(n_obs = n()) %>% 
  mutate(race_cat = ifelse(is.na(race_cat), "missing",
                           as.character(race_cat)),
         total = sum(n_obs),
         prop = round(n_obs/total,2)) %>% 
  dplyr::select(race_cat, n_obs, prop, total)

race_tab
```

## PM~2.5~ Distribution

I'm going to read in the GWR estimates Kate produced for Colorado.

Starting by reading in Western US grid and limiting to same study extent as the previous WRF-GRID. I've also plotted the 15x15 km grid (red) over the 3x3 km grid (blue) we have for the Colorado morbidity data. The grids will have to be linked in the future, but I don't think it will be too tricky. I think you could just take the centroids of the smaller grid and do a spatial join to the larger grid.

```{r wus_grid}
# read grid sf
wus_grid_sf <- st_read(paste0("../../meta_wildfire/data/shapefile/",
  "2015-wus_grid"), layer="wus_grid2015-ko")

# set projection
st_crs(wus_grid_sf) <- wgs
# clipping bounding box same as other grid
study_bbox <- st_bbox(c(xmin=-105.3, xmax=-104.5, ymax=41, ymin = 38))
# clip to bbox extent
gwr_grid_clip <- bbox_clip(wus_grid_sf, study_bbox)

ggplot(gwr_grid_clip) +
  geom_sf(color = "red", fill = NA) +
  geom_sf(data=study_grid, color = "blue", fill=NA) +
  theme_minimal()
```

Extracting PM estimates from .nc file.

```{r nc_connection}
# open connection to netcdf in meta-wildfire repo
nc_path <- paste0("../../meta_wildfire/data/smoke/2015-smoke_pm_wus.nc")
smk_2015_nc <- nc_open(nc_path)
# print summary
smk_2015_nc
```

Extracting grid points (matrix) and setting as vectors. When using the as.vector function on matrix values from a netcdf, it will go row by row (left to right, then top to bottom).

```{r extract_location_values}
# extract longitude as a vector
lon <- ncvar_get(smk_2015_nc, "glons")
# extract latitude as a vector
lat <- ncvar_get(smk_2015_nc, "glats")
# extract date as a vector (format yyyymmdd)
date <- ncvar_get(smk_2015_nc, "dates")

# vector of lat and lon
lon_vec <- as.vector(lon)
lat_vec <- as.vector(lat)
# id sequence
id <- 1:20838

# bind together and plot to see where grids are
id_location <- as_data_frame(cbind(id, lon_vec, lat_vec))

# 138 columns, 151 rows
# recreating a matrix that matches values
grid_id <- t(apply(matrix(20838:1, nrow=151, ncol=138, byrow = T),1,rev))
# extracting values as a vector that matches grid values in netcdf
cell_id <- as.vector(grid_id)

# WRFGRID_ID matrix based on cell location
wrf_id_mat <- apply(matrix(1:20838, nrow=151, ncol=138), 2, rev)
# vector of WRFGRID_ID
WRFGRID_ID <- as.vector(wrf_id_mat)

# join WRFGRID_ID values to lat lon coordinates
wrf_id_key <- as_data_frame(cbind(cell_id, WRFGRID_ID)) %>% 
  rename(id = cell_id) %>% 
  left_join(id_location, by = "id") %>% 
  rename(lat = lat_vec, lon = lon_vec) %>% 
  mutate(combine_id = paste(id, WRFGRID_ID, sep = ":"))

# subset nc ids by WRFID grid
study_area_id_key <- wrf_id_key %>% 
  filter(WRFGRID_ID %in% gwr_grid_clip$WRFGRID_ID)
```

Extracting PM~2.5~ values for netcdf file. This is for the entire W. US, which is not super efficient, but it's probably faster to do this since I've already QCed this rather than figure out how to find the specific grids from scratch.
```{r pm_vals}
# extract krig_pm and background_pm from nc file ----
gwr_pm <- ncvar_get(smk_2015_nc, "gwrPM")
krig_pm <- ncvar_get(smk_2015_nc, "kPM")
background_pm <- ncvar_get(smk_2015_nc, "gwr_background")

# create cell x date matrices ----
# apply function to extract pm values for each cell for each date as a vector
krig_pm_mat <- apply(krig_pm, 3, as.vector)
gwr_pm_mat <- apply(gwr_pm, 3, as.vector)
# background has only one value
background_pm_vec <- as.vector(background_pm) 

# set values <0 to 0
krig_pm_mat <- ifelse(krig_pm_mat < 0, 0, krig_pm_mat)
gwr_pm_mat <- ifelse(gwr_pm_mat <0, 0, gwr_pm_mat)

# create smoke estimates -----
krig_pm_smk_mat <- krig_pm_mat - background_pm_vec
gwr_pm_smk_mat <- gwr_pm_mat - background_pm_vec
# set values lower than 0 to 0
krig_pm_smk_mat <- ifelse(krig_pm_smk_mat < 0, 0, krig_pm_smk_mat)
gwr_pm_smk_mat <- ifelse(gwr_pm_smk_mat < 0, 0, gwr_pm_smk_mat)

# assign WRFGRID IDs as rownames
# extract vector of arranged WRFGRID IDs that match cell IDs for rownames
WRFGRID <- arrange(wrf_id_key, id)$WRFGRID_ID
# extract dates for column names
date_vec <- paste0("date", date)

# create list of pm values
pm_list <- list(krig_pm_mat, krig_pm_smk_mat, gwr_pm_mat, gwr_pm_smk_mat)
# assign names to pm_list
names(pm_list) <- c("krig", "krig_smk", "gwr", "gwr_smk")

study_area_wrfid <- study_area_id_key$WRFGRID_ID
# lapply row and column names and sort by WRFGRID Cell, then extract by 
# wrf ids in study area
pm_list <- lapply(pm_list, function(x){
  colnames(x) <- date_vec
  rownames(x) <- WRFGRID
  x <- x[order(as.numeric(row.names(x))),]
  x <- x[study_area_wrfid, ]
})

# subsetting PM GWR as a dataframe and preparing for distribution
gwr_pm <- data.frame(pm_list[[3]]) %>% 
  cbind(study_area_wrfid, .) %>% 
  rename(WRFGRID_ID = study_area_wrfid) %>% 
  gather(key = "date", value = "gwr_pm", -WRFGRID_ID) %>% 
  mutate(date = as.Date(stringr::str_sub(date, start=5), 
                            format = "%Y%m%d"))

gwr_smk <- data.frame(pm_list[[4]]) %>% 
  cbind(study_area_wrfid, .) %>% 
  rename(WRFGRID_ID = study_area_wrfid) %>% 
  gather(key = "date", value = "gwr_pm_smk", -WRFGRID_ID) %>% 
  mutate(date = as.Date(stringr::str_sub(date, start=5), 
                            format = "%Y%m%d")) %>% 
  # join pm values
  left_join(gwr_pm, by = c("WRFGRID_ID", "date")) %>% 
  mutate(gwr_smk0 = ifelse(gwr_pm_smk > 0, 1, 0),
         gwr_smk5 = ifelse(gwr_pm_smk > 5, 1, 0),
         gwr_smk10 = ifelse(gwr_pm_smk > 10, 1, 0),
         gwr_smk15 = ifelse(gwr_pm_smk > 15, 1, 0),
         gwr_smk20 = ifelse(gwr_pm_smk > 20, 1, 0)) %>% 
  gather(key = "smoke_cat", value = "smoke_indicator", 
         -c(WRFGRID_ID, date, gwr_pm_smk, gwr_pm))  %>% 
  mutate(smoke_indicator = as.factor(smoke_indicator),
         smoke_cat = forcats::fct_relevel(smoke_cat, 
            c("gwr_smk0", "gwr_smk5", "gwr_smk10", "gwr_smk15",
              "gwr_smk20")))
```

### PM~2.5~ Distributions by Smoke Cut Point
Distributions of PM~2.5~ estimate by GWR by smoke indicators for 5 different smoke cutoff points. 
```{r pm_smk_dist}
dist_plot <- ggplot(gwr_smk, 
  aes(x=gwr_pm, group=smoke_indicator, fill=smoke_indicator)) +
  geom_density(alpha = 0.6) +
  facet_wrap(~smoke_cat) +
  theme_minimal()

dist_plot
```
For my binary smoke classifiers, I think the cutoffs at 0 and 5 are prone to exposure misclassification. Cutpoints 15 and 20 may be the most specific.

### PM~2.5~ Summary Statistics 
The variable smoke_cat is the binary cutoff, smoke_indicator indicates possible smoke where based on gwr_smk = (gwr_pm - background), n_obs is the number of observations for each smoke indicator for the grids in the study area, followed by minimum, maximum, median, and mean GWR PM~2.5~ values (not the gwr_smk estimates).
```{r pm_sum_stat}
pm_summary <- gwr_smk %>% 
  group_by(smoke_cat, smoke_indicator) %>% 
  summarise(n_obs = n(), min_pm = min(gwr_pm), max_pm = max(gwr_pm), 
            med_pm = median(gwr_pm), mean_pm = mean(gwr_pm))
  
pm_summary
```


